### 2022-01-03 README.Rmd
map, temp logs, pca, nmds figs

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![](hou.jpg)

# Picos

Proyecto Picos.

## Datos pendientes

## Análisis pendientes

### 2022-01-03 src/dca.R

@@ -1,77 +1,0 @@
-read.csv("data/ibutton-plot-species.csv") %>%
-  group_by(Taxon) %>% 
-  tally %>%
-  arrange(-n) %>%
-  data.frame %>%
-  filter(n >2) %>%
-  pull(Taxon) -> spp
-
-read.csv("data/ibutton-plot-header.csv") %>%
-  select(Plot, Site) %>%
-  merge(read.csv("data/ibutton-plot-species.csv")) %>%
-  # filter(Site == "Los Boches") %>%
-  filter(Plot != "D19") %>%
-  select(-Site) %>%
-  spread(Taxon, Cover, fill = 0) %>%
-  arrange(Plot) %>%
-  column_to_rownames(var = "Plot") -> df4
-
-read.csv("data/ibutton-plot-header.csv") %>%
-  select(Plot, Site) %>%
-  merge(read.csv("data/ibutton-plot-species.csv")) %>%
-  # filter(Site == "Los Boches") %>%
-  filter(Plot != "D19") %>%
-  select(Plot) %>% unique %>%
-  arrange(Plot) %>%
-  merge(micro) %>%
-  select(-c(Site, Plot, `Annual\ntemperature`, `Annual range`, `Diurnal range`)) -> env
-
-vegan::decorana(df4) -> dca1
-
-vegan::envfit(dca1, env) -> ef1
-
-ef1$vectors$arrows %>%
-  data.frame() %>%
-  select(DCA1, DCA2) %>%
-  rownames_to_column(var = "Variable") -> vars
-
-ef1$vectors
-
-vegan::scores(dca1, "species") %>%
-  data.frame() %>%
-  rownames_to_column("Species") %>%
-  filter(Species %in% spp) -> 
-  sppscores
-
-vegan::scores(dca1, "sites") %>%
-  data.frame %>%
-  rownames_to_column(var = "Plot") %>%
-  merge(groups) -> df5
-
-ggplot(df5, aes(x = DCA1, y = DCA2)) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  geom_vline(xintercept = 0, linetype = "dashed") +
-  geom_segment(data = vars, aes(x = 0, y = 0, xend = 3*DCA1, yend = 3*DCA2)) +
-  ggthemes::theme_tufte() + 
-  ggrepel::geom_text_repel(data = sppscores, aes(label = Species), fontface = "italic", size = 3) + 
[…]

### 2022-01-03 src/figure-map.R

@@ -1,107 +1,0 @@
-library(tidyverse); library(raster)
-
-### Get map of Spain
-
-spaingeo <- getData("GADM", country = "spain", level = 1)
-spaindf <- fortify(spaingeo)
-
-spainalt <- getData("alt", country = "spain", level = 1)
-altdf  <- data.frame(rasterToPoints(spainalt))
-colnames(altdf) = c("lon", "lat", "alt")
-altdf %>% filter(alt < 2650) -> altdf
-
-### Site coordinates
-
-rbind(
-read.csv("data/focal-plot-header.csv") %>%
-  filter(Site %in% c("Hoyo Sin Tierra", 
-                     "Los Boches", 
-                     "Hou Sin Tierri",
-                     "Los Cazadores")) %>%
-  group_by(Site) %>%
-  summarise(long = mean(Longitude), lat = mean(Latitude)) %>%
-  mutate(Plot = "Focal"),
-read.csv("data/ibutton-plot-header.csv") %>%
-  group_by(Site, Plot) %>%
-  summarise(long = mean(Longitude), lat = mean(Latitude)) %>%
-  mutate(Plot = "Transect")) %>%
-  mutate(Site = fct_relevel(Site, "El Cable", "Hoyo Sin Tierra", "Cabaña Verónica",
-                            "Horcados Rojos", "Los Boches", "Hou Sin Tierri",
-                            "Los Cazadores", "Urriellu")) %>%
-  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> header
-
-### Draw map of Spain
-
-header %>%
-  filter(Plot == "Focal") %>%
-  ggplot(aes(long, lat)) +
-  geom_tile(data = altdf, aes(x = lon, y = lat, fill = alt)) +
-  scale_fill_gradient(name = "Elevation (m asl)", low = "grey96", high = "chocolate4",
-                      breaks = c(0, 2600)) +
-  geom_blank(data = spaindf, aes(x = long, y = lat)) +
-  geom_map(data = spaindf, map = spaindf,
-           aes(group = group, map_id = id),
-           fill = NA, color = "black", size = .3) +
-  ggrepel::geom_label_repel(aes(label = Site), box.padding = .6, size = 4,
-                            family = "serif", fontface = "italic") +
-  labs(x = "Longitude (º)", y = "Latitude (º)") +
-  geom_point(size = 1, color = "black", fill = "#9a02db", shape = 22) +
[…]

### 2022-01-03 src/figure-nmds-species.R

@@ -1,0 +1,136 @@
+library(tidyverse)
+
+### Vegetation data from the spatial survey
+
+read.csv("data/spatial-survey-species.csv") %>%
+  filter(Plot != "A03") %>%
+  filter(Plot != "D19") %>%
+  spread(Taxon, Cover, fill = 0) %>%
+  arrange(Plot) %>%
+  column_to_rownames(var = "Plot") -> species
+
+### Do NMDS
+
+species %>%
+  vegan::metaMDS(trymax = 100000, k = 2) -> 
+  nmds # Ordination output
+
+### Site scores
+
+vegan::scores(nmds, "sites") %>%
+  data.frame() %>%
+  rownames_to_column("Plot") %>%
+  rename(Dim.1 = NMDS1, Dim.2 = NMDS2) %>%
+  merge(read.csv("data/spatial-survey-header.csv")) %>%
+  dplyr::select(Plot, Site, Dim.1, Dim.2) %>% 
+  mutate(Site = fct_relevel(Site,
+                            "Los Cazadores", "Hou Sin Tierri",
+                            "Los Boches", "Hoyo Sin Tierra")) %>%
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> sitescores
+
+### Species scores
+
+read.csv("data/spatial-survey-species.csv") %>%
+  group_by(Taxon) %>%
+  tally %>%
+  filter(n > 80 * .2) %>%
+  pull(Taxon) -> sppfreq
+
+vegan::scores(nmds, "species") %>%
+  data.frame() %>%
+  rownames_to_column("Taxon") %>%
+  # filter(Taxon %in% sppfreq) %>%
+  rename(Dim.1 = NMDS1, Dim.2 = NMDS2) %>%
+  mutate(Species = gsub(" gr\\.", "", Taxon)) %>%
+  mutate(Species = gsub("apenninum subsp. ", "", Species)) %>%
+  separate(Species, into = c("G", "S")) %>%
+  mutate(Species = paste(substr(G, 1, 3), substr(S, 1, 3), sep = "")) %>%
+  dplyr::select(Taxon, Species, Dim.1, Dim.2)-> sppscores
+
+### Environmental scores
+
+read.csv("data/spatial-survey-temperatures.csv") %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
[…]

### 2022-01-03 src/figure-pca-temperatures.R

@@ -1,0 +1,71 @@
+library(tidyverse)
+
+### Spatial survey data
+
+read.csv("data/spatial-survey-temperatures.csv") %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
+  merge(read.csv("data/spatial-survey-header.csv")) %>%
+  group_by(Plot, Site, Day = lubridate::floor_date(Time, "day")) %>%
+  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
+  mutate(Snow = ifelse(X < 0.5 & N > -0.5, 1, 0)) %>% # Day is snow day or not
+  mutate(FreezeThaw = ifelse(X > 0.5 & N < -0.5, 1, 0)) %>% # Day with freeze-thaw cycles
+  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
+  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
+  group_by(Plot, Site, Month = lubridate::floor_date(Day, "month")) %>%
+  summarise(T = mean(T), X = mean(X), N = mean(N), # Daily mean, max, min
+            Snow = sum(Snow), # Snow days per month
+            FreezeThaw = sum(FreezeThaw), # Freeze-thaw days per month
+            FDD = sum(FDD), # FDD per month
+            GDD = sum(GDD)) %>% # GDD per month
+  group_by(Plot, Site) %>%
+  summarise(`Annual\ntemperature` = mean(T), # Annual Mean Temperature
+            `Diurnal range` = mean(X - N), # Mean Diurnal Range (Mean of monthly (max temp - min temp))
+            `Annual range` = max(X) - min(N), # Temperature Annual Range (BIO5-BIO6)
+            `Freezing\ndegrees-day` = sum(FDD), # FDD per year
+            `Growing\ndegrees-day` = sum(GDD)) -> # GDD per year
+  micro
+
+### PCA
+
+micro[, 3:7] %>%
+  FactoMineR::PCA() -> pca
+
+cbind((micro %>%  dplyr::select(Plot, Site)), data.frame(pca$ind$coord[, 1:2])) %>%
+  mutate(Site = fct_relevel(Site,
+                            "Los Cazadores", "Hou Sin Tierri",
[…]

### 2022-01-03 src/figure-spp-changes.R

@@ -1,0 +1,82 @@
+### NMDS species scores
+
+read.csv("data/spatial-survey-species.csv") %>%
+  filter(Plot != "A03") %>%
+  filter(Plot != "D19") %>%
+  spread(Taxon, Cover, fill = 0) %>%
+  arrange(Plot) %>%
+  column_to_rownames(var = "Plot") %>%
+  vegan::metaMDS(trymax = 100000, k = 2) %>%
+  vegan::scores(., "species") %>%
+  data.frame() %>%
+  rownames_to_column("Taxon") %>%
+  rename(Dim.1 = NMDS1, Dim.2 = NMDS2) %>%
+  mutate(Species = gsub(" gr\\.", "", Taxon)) %>%
+  mutate(Species = gsub("apenninum subsp. ", "", Species)) %>%
+  separate(Species, into = c("G", "S")) %>%
+  mutate(Species = paste(substr(G, 1, 3), substr(S, 1, 3), sep = "")) %>%
+  dplyr::select(Taxon, Species, Dim.1, Dim.2) -> sppscores
+
+### Species changes
+
+read.csv("data/temporal-survey-matrices.csv") %>%
+  group_by(Plot, Taxon) %>%
+  summarise(Cover = sum(Present)) %>%
+  merge(read.csv("data/species-checklist.csv"), all.x = TRUE) %>%
+  filter(Lifeform != "Therophyte") %>%
+  separate(Plot, into = c("Plot", "Year"), sep = "-") %>%
+  mutate(Year = fct_recode(Year, "09" = "08")) %>%
+  group_by(Taxon, Year) %>%
+  summarise(Cover = sum(Cover)) %>%
+  spread(Year, Cover) %>%
+  mutate(Change = `19` - `09`) %>%
+  mutate(ChangeP = `19` / `09`) %>%
+  arrange(Change) %>%
+  na.omit %>%
+  merge(sppscores) -> changespp
+
+### Plot decreasing
+
+changespp %>%
+  filter(Change < -10) %>%
+  ggplot(aes(Dim.1, Dim.2)) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  geom_vline(xintercept = 0, linetype = "dashed") +
+  geom_segment(data = envvars, aes(x = 0, y = 0, xend = .3*Dim.1, yend = .3*Dim.2)) +
+  geom_point(aes(size = abs(Change), fill = Change), shape = 21, show.legend = FALSE) +
[…]

### 2022-01-03 src/figure-survey-sites.R

@@ -1,0 +1,150 @@
+library(tidyverse); library(raster); library(tibbletime); library(lubridate)
+
+### Raster with elevation for Picos area
+
+raster::raster("../#data/maps/picos/DEM/DEM.TIF") %>% # Map files are in my home drive
+  rasterToPoints() %>%
+  as.data.frame() %>%
+  rename(lon = x, lat = y, alt = DEM) -> dem 
+
+### Site coordinates
+
+read.csv("data/temporal-survey-header.csv") %>%
+  group_by(Site) %>%
+  summarise(long = mean(ED5030TX), lat = mean(ED5030TY)) %>%
+  arrange(-lat) %>%
+  rownames_to_column(var = "Label") -> tPlots
+
+### Panel A - Map of Europe
+
+rgdal::readOGR(dsn = "../#data/maps/WWF", 
+               layer = "wwf_terr_ecos") -> Ecoregions # Map files are in my home drive
+rownames(Ecoregions@data) -> Ecoregions@data$id
+fortify(Ecoregions, region = "id") -> Ecoregions.points 
+plyr::join(Ecoregions.points, Ecoregions@data, by = "id") %>%
+  inner_join(read.csv("../#data/maps/Biomes.csv"), by = "BIOME") -> land
+
+read.csv("data/temporal-survey-header.csv") %>%
+  summarise(long = mean(Longitude), lat = mean(Latitude)) %>%
+  ggplot(aes(long, lat)) +
+  geom_polygon(data = land, aes(x = long, y = lat, group = group), 
+               color = NA, fill = "gainsboro", size = 0.25, show.legend = FALSE) +
+  ggstar::geom_star(size = 4, color = "black", fill = "#9a02db", starshape = 1) +
+  labs(x = "Longitude (º)", y = "Latitude (º)") +
+  coord_cartesian(xlim = c(-9, 21), ylim = c(36.5, 58.5)) +
+  ggthemes::theme_tufte() +
+  labs(title = "(A) Study area") +
+  theme(panel.background = element_rect(color = NA, fill = "grey96"),
+        legend.position = "bottom",
+        legend.direction = "horizontal",
+        legend.background = element_blank(),
+        legend.box.background = element_rect(colour = NA),
+        legend.title = element_text(size = 10),
+        legend.text = element_text(size = 8),
+        legend.justification = "left",
+        legend.margin = margin(0, 0, 0, 0),
[…]

### 2022-01-03 src/figure-temperature-changes.R

@@ -1,0 +1,57 @@
+library(tidyverse); library(urca)
+
+### Decompose time series
+
+decomposer <- function(x) 
+{
+  x %>% 
+    pull(Temperature) %>%
+    ts(start = c(2009, 1), frequency = 24 * 365) -> timeseries
+  
+  stl(timeseries, s.window = "period") -> fit
+  
+  cbind(x$Time, as.data.frame(timeseries), as.data.frame(fit$time.series)) -> df
+  
+  colnames(df) <- c("Time", "Temperature", "Seasonal", "Trend", "Remainder")
+  
+  df
+}
+
+read.csv("data/temporal-survey-temperatures.csv") %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
+  group_by(Site) %>%
+  do(decomposer(.)) %>%
+  mutate(Site = as.factor(Site)) %>%
+  mutate(Site = fct_relevel(Site,
+                            "Los Cazadores", "Hou Sin Tierri",
+                            "Los Boches", "Hoyo Sin Tierra")) %>%
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> trends
+
+### Plot trends
+
+trends %>%
+  ggplot(aes(Time, Trend)) + 
+  facet_wrap(~ Site, nrow = 1) + 
+  geom_line(color = "dodgerblue") +
+  geom_smooth(method = "lm", color = "dodgerblue4") +
+  xlab("Time (1-h recording inverval)") +
+  ylab("Trend (ºC)") +
+  ggthemes::theme_tufte() +
+  theme(legend.position = "none", 
+        legend.direction = "vertical",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 16, face = "italic"), 
+        panel.background = element_rect(color = "grey96", fill = "grey96"),
+        strip.text = element_text(size = 12),
+        axis.title = element_text(size = 12),
+        axis.text.x = element_text(size = 8, color = "black"),
+        axis.text.y = element_text(size = 12, color = "black")) -> f1; f1
+
+### Save figure
+
+ggsave(f1, file = "results/figures/temperature-changes.png", 
+       path = NULL, scale = 1, width = 182, height = 182/3, units = "mm", dpi = 600)
+# ggsave(f1, file = "results/figures/temperature-changes.tiff", device = grDevices::tiff, 
[…]

### 2022-01-03 src/figure-temperature-logs.R

@@ -1,0 +1,65 @@
+library(tidyverse)
+Sys.setlocale("LC_TIME", "English")
+
+### Geoprecision logs
+
+read.csv("data/temporal-survey-temperatures.csv") %>% 
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
+  mutate(Site = fct_relevel(Site,
+                            "Los Cazadores", "Hou Sin Tierri",
+                            "Los Boches", "Hoyo Sin Tierra")) %>%
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) %>%
+  ggplot(aes(Time, Temperature)) + 
+  facet_wrap(~ Site, nrow = 1) + 
+  geom_line(color = "red") +
+  labs(title = "(A) Temperature logs from the temporal survey (2009-2018)") +
+  xlab("Time (1-h recording inverval)") +
+  ylab("Temperature (ºC)") +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  ggthemes::theme_tufte() +
+  theme(legend.position = "none", 
+        legend.direction = "vertical",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 16, face = "italic"), 
+        panel.background = element_rect(color = "grey96", fill = "grey96"),
+        strip.text = element_text(size = 12),
+        axis.title = element_text(size = 12),
+        axis.text.x = element_text(size = 8, color = "black"),
+        axis.text.y = element_text(size = 12, color = "black")) -> f1; f1
+
+### iButton logs
+
+read.csv("data/spatial-survey-temperatures.csv") %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
+  merge(read.csv("data/spatial-survey-header.csv")) %>%
+  mutate(Site = fct_relevel(Site,
+                            "Los Cazadores", "Hou Sin Tierri",
+                            "Los Boches", "Hoyo Sin Tierra")) %>%
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) %>%
+  ggplot(aes(Time, Temperature, color = Plot)) + 
+  facet_wrap(~ Site, nrow = 1) + 
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  geom_line() +
+  labs(title = "(B) Temperature logs from the spatial survey (20 iButtons per site)") +
[…]

### 2022-01-03 src/functional-traits.R

@@ -1,9 +1,0 @@
-library(tidyverse)
-
-read.csv("data/ibutton-plot-species.csv") %>%
-  merge(read.csv("data/species-traits.csv"), all.x = TRUE) %>%
-  mutate(Traited = ifelse(is.na(SLA), "No", "Yes")) %>%
-  group_by(Plot, Traited) %>%
-  summarise(Cover = sum(Cover)) %>%
-  spread(Traited, Cover) %>%
-  mutate(Coverage = Yes / (No + Yes)) %>% filter(Coverage > .8)

### 2022-01-03 src/ibutton data fig.R

@@ -1,28 +1,0 @@
-read.csv("data/ibutton-plot-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
-  merge(read.csv("data/ibutton-plot-header.csv")) %>%
-  mutate(Site = fct_relevel(Site, "El Cable", "Hoyo Sin Tierra", "Cabaña Verónica",
-                            "Horcados Rojos", "Los Boches", "Hou Sin Tierri",
-                            "Los Cazadores", "Urriellu")) %>%
-  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) %>%
-  ggplot(aes(Time, Temperature, color = Plot)) + 
-  facet_wrap(~ Site, nrow = 2) + 
-  geom_line() +
-  labs(title = "Microspatial variation in soil temperature (n = 20 iButtons per site)") +
-  xlab("Time (4 h)") +
-  ylab("Temperature (ºC)") +
-  ggthemes::theme_tufte() +
-  theme(legend.position = "none", 
-        legend.direction = "vertical",
-        legend.title = element_blank(),
-        legend.text = element_text(size = 16, face = "italic"), 
-        strip.text = element_text(size = 22), 
-        panel.background = element_rect(color = "grey96", fill = "grey96"),
-        title = element_text(size = 24, color = "black", face = "bold"),
-        axis.title.y = element_text(size = 22, color = "black"),
-        axis.title.x = element_text(size = 22, color = "black"),
-        axis.text.y = element_text(size = 14, color = "black"),
-        axis.text.x = element_text(size = 14, color = "black")) -> f1; f1
-
-ggsave(f1, file = "results/ibutton data.png", 
-       path = NULL, scale = 1, width = 330, height = 185, units = "mm", dpi = 600)

### 2022-01-03 src/pca-ibuttons.R

@@ -1,74 +1,0 @@
-read.csv("data/ibutton-plot-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
-  merge(read.csv("data/ibutton-plot-header.csv")) %>%
-  group_by(Plot, Site, Day = lubridate::floor_date(Time, "day")) %>%
-  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
-  mutate(Snow = ifelse(X < 0.5 & N > -0.5, 1, 0)) %>% # Day is snow day or not
-  mutate(FreezeThaw = ifelse(X > 0.5 & N < -0.5, 1, 0)) %>% # Day with freeze-thaw cycles
-  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
-  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
-  group_by(Plot, Site, Month = lubridate::floor_date(Day, "month")) %>%
-  summarise(T = mean(T), X = mean(X), N = mean(N), # Daily mean, max, min
-            Snow = sum(Snow), # Snow days per month
-            FreezeThaw = sum(FreezeThaw), # Freeze-thaw days per month
-            FDD = sum(FDD), # FDD per month
-            GDD = sum(GDD)) %>% # GDD per month
-  group_by(Plot, Site) %>%
-  summarise(`Annual\ntemperature` = mean(T), # Annual Mean Temperature
-            `Diurnal range` = mean(X - N), # Mean Diurnal Range (Mean of monthly (max temp - min temp))
-            # bio04 = sd(T) * 100, # Temperature Seasonality (standard deviation ×100)
-            # bio05 = max(X), # Max Temperature of Warmest Month
-            # bio06 = min(N), # Min Temperature of Coldest Month
-            `Annual range` = max(X) - min(N), # Temperature Annual Range (BIO5-BIO6)
-            # bio03 = (bio02 / bio07) * 100, # Isothermality (BIO2/BIO7) (×100)
-            # Snow = sum(Snow), # Days of snow per year
-            # FreezeThaw = sum(FreezeThaw), # Days with freeze-thaw cycles per year
-            `Freezing\ndegrees-day` = sum(FDD), # FDD per year
[…]

### 2022-01-03 src/time-series.R

@@ -1,192 +1,0 @@
-library(tidyverse); library(urca)
-
-read.csv("data/focal-plot-temperatures.csv") -> temperatures
-
-# # El Cable
-# 
-# temperatures %>% 
-#   filter(Site == "El Cable") %>% 
-#   pull(Temperature) %>%
-#   ts(start = c(2009, 1), frequency = 24 * 365) -> # Convert to time series object
-#   timeseries
-# 
-# fit <- stl(timeseries, s.window = "period") # Better than `decompose` https://stats.stackexchange.com/questions/85987/which-is-better-stl-or-decompose
-# plot(fit)
-# 
-# tseries::adf.test(timeseries) # Time series IS stationary
-# Kendall::SeasonalMannKendall(timeseries) # There is a trend in the data https://www.statology.org/mann-kendall-trend-test-r/
-# 
-# # Hoyo Sin Tierra
-# 
-# temperatures %>% 
-#   filter(Site == "Hoyo Sin Tierra") %>% 
-#   pull(Temperature) %>%
-#   ts(start = c(2009, 1), frequency = 24 * 365) -> # Convert to time series object
-#   timeseries
-# 
-# fit <- stl(timeseries, s.window = "period") # Better than `decompose` https://stats.stackexchange.com/questions/85987/which-is-better-stl-or-decompose
-# plot(fit)
-# 
-# tseries::adf.test(timeseries) # Time series IS stationary
-# Kendall::SeasonalMannKendall(timeseries) # There is a trend in the data https://www.statology.org/mann-kendall-trend-test-r/
-# 
-# # Cabaña Verónica
-# 
-# temperatures %>% 
-#   filter(Site == "Cabaña Verónica") %>% 
-#   pull(Temperature) %>%
-#   ts(start = c(2009, 1), frequency = 24 * 365) -> # Convert to time series object
-#   timeseries
-# 
-# fit <- stl(timeseries, s.window = "period") # Better than `decompose` https://stats.stackexchange.com/questions/85987/which-is-better-stl-or-decompose
-# plot(fit)
-# 
-# tseries::adf.test(timeseries) # Time series IS stationary
-# Kendall::SeasonalMannKendall(timeseries) # There is a trend in the data https://www.statology.org/mann-kendall-trend-test-r/
-# 
-# # Horcados Rojos
-# 
-# temperatures %>% 
-#   filter(Site == "Horcados Rojos") %>% 
-#   pull(Temperature) %>%
[…]

### 2022-01-03 src/vegetation-changes2.R

@@ -1,226 +1,0 @@
-library(tidyverse)
-
-read.csv("data/focal-plot-matrices.csv") %>%
-  merge(read.csv("data/focal-plot-header.csv")) %>%
-  filter(Site %in% c("Hou Sin Tierri", "Hoyo Sin Tierra", "Los Boches", "Los Cazadores")) %>%
-  select(Plot, Cell, Taxon, Present, Site) %>%
-  merge(read.csv("data/Grid.csv")) %>%
-  separate(Plot, into = c("Plot", "Year"), sep = "-") %>%
-  mutate(Year = fct_recode(Year, "09" = "08")) %>%
-  mutate(Year = fct_recode(Year, "2009" = "09")) %>%
-  mutate(Year = fct_recode(Year, "2019" = "19")) %>%
-  select(Site, Plot, Year, Taxon, Row, Column, Present) %>%
-  arrange(Site, Plot, Year, Taxon, Row, Column, Present) %>%
-  filter(Taxon == "Helianthemum apenninum subsp. urrielense") %>%
-  # filter(Plot == "HI2") %>%
-  ggplot(aes(Column, Row, fill = Present)) +
-  facet_grid(Year ~ Plot) +
-  geom_tile(fill = "limegreen") +
-  labs(title = "10-year changes", subtitle = "Helianthemum apenninum subsp. urrielense") +
-  xlab("Rows") +
-  ylab("Columns") +
-  ggthemes::theme_tufte() +
-  theme(legend.position = "none", 
-        legend.direction = "vertical",
-        legend.title = element_blank(),
-        legend.text = element_text(size = 16, face = "italic"), 
-        strip.text = element_text(size = 22), 
-        panel.background = element_rect(color = "grey96", fill = "grey96"),
-        title = element_text(size = 24, color = "black", face = "bold"),
-        plot.subtitle = element_text(size = 20, color = "black", face = "italic"),
-        axis.title.y = element_blank(),
-        axis.title.x = element_blank(),
-        axis.text.y = element_blank(),
-        axis.text.x =element_blank(),
-        axis.ticks = element_blank()) -> urrielense; urrielense
-
-ggsave(urrielense, file = "results/urrielense.png", 
-       path = NULL, scale = 1,width = (330/3)*2, height = 185, units = "mm", dpi = 600) 
-
-library(tidyverse)
-
-read.csv("data/focal-plot-matrices.csv") %>%
[…]

### 2022-01-05 src/figure-nmds-species.R
figures v1

@@ -62,7 +62,7 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
   summarise(FDD = sum(FDD), # FDD per month
             GDD = sum(GDD)) %>% # GDD per month
   group_by(Plot, Site) %>%
-  summarise(`Freezing\ndegrees-day` = sum(FDD), # FDD per year
+  summarise(`Freezing\ndegrees-day` = abs(sum(FDD)), # FDD per year
             `Growing\ndegrees-day` = sum(GDD)) %>%
   group_by() %>%
   arrange(Plot) %>%
@@ -83,54 +83,32 @@ ef1$vectors$arrows %>%
 ggplot(sitescores, aes(x = Dim.1, y = Dim.2)) +
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_vline(xintercept = 0, linetype = "dashed") +
-  geom_segment(data = envvars, aes(x = 0, y = 0, xend = .7*Dim.1, yend = .7*Dim.2)) +
-  geom_point(aes(fill = Site), size = 4, shape = 22) +
-  geom_label(data = envvars, aes(x = .7*Dim.1, y = .7*Dim.2, label = Variable),  show.legend = FALSE, size = 2.5) +
-  ggthemes::theme_tufte() + 
-  guides(fill = guide_legend(ncol = 2)) +
-  theme(legend.position = c(.5, .1), 
-        legend.title = element_blank(),
-        panel.background = element_rect(color = "grey96", fill = "grey96"),
-        axis.title = element_text(size = 10),
-        axis.text = element_text(size = 10, color = "black")) +
-  coord_cartesian(xlim = c(-1, 1), ylim = c(-1, 1)) +
-  labs(title = "(A) Spatial survey - sites") +
-  scale_x_continuous(name = "NMDS1") + 
-  scale_y_continuous(name = "NMDS2") +
-  scale_color_manual(values = c("goldenrod", "forestgreen",  "royalblue", "darkorchid")) +
-  scale_fill_manual(values = c("goldenrod", "forestgreen",  "royalblue", "darkorchid")) -> f1; f1
-
-### Plot NMDS sites
-
-ggplot(sitescores, aes(x = Dim.1, y = Dim.2)) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  geom_vline(xintercept = 0, linetype = "dashed") +
-  ggrepel::geom_text_repel(data = sppscores, aes(label = Species), fontface = "italic", size = 2, segment.color = "transparent") +
[…]

### 2022-01-05 src/figure-pca-temperatures.R

@@ -21,7 +21,7 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
   summarise(`Annual\ntemperature` = mean(T), # Annual Mean Temperature
             `Diurnal range` = mean(X - N), # Mean Diurnal Range (Mean of monthly (max temp - min temp))
             `Annual range` = max(X) - min(N), # Temperature Annual Range (BIO5-BIO6)
-            `Freezing\ndegrees-day` = sum(FDD), # FDD per year
+            `Freezing\ndegrees-day` = abs(sum(FDD)), # FDD per year
             `Growing\ndegrees-day` = sum(GDD)) -> # GDD per year
   micro
 
@@ -48,18 +48,19 @@ ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
   geom_vline(xintercept = 0, linetype = "dashed") +
   geom_segment(data = pcaVars, aes(x = 0, y = 0, xend = 3*Dim.1, yend = 3*Dim.2)) +
   ggthemes::theme_tufte() + 
-  geom_point(aes(fill = Site), size = 4, shape = 22) +
+  geom_point(aes(fill = Site), size = 6, shape = 22) +
   theme(legend.position = "top", 
         legend.title = element_blank(),
+        # legend.text = element_text(size = 12, color = "black"),
         panel.background = element_rect(color = "grey96", fill = "grey96"),
-        axis.title = element_text(size = 10),
-        axis.text = element_text(size = 10, color = "black"),
+        axis.title = element_text(size = 12),
+        axis.text = element_text(size = 12, color = "black"),
         plot.margin = unit(c(0, 0.1, 0, 0), "cm")) +
   scale_x_continuous(name = paste("Axis 1 (", round(pca$eig[1, 2], 0),
                                   "% variance explained)", sep = "")) + 
   scale_y_continuous(name = paste("Axis 2 (", round(pca$eig[2, 2], 0), 
                                   "% variance explained)", sep = "")) +
-  geom_label(data = pcaVars, aes(x = 3*Dim.1, y = 3*Dim.2, label = Variable),  show.legend = FALSE, size = 2.5) +
+  geom_label(data = pcaVars, aes(x = 3*Dim.1, y = 3*Dim.2, label = Variable),  show.legend = FALSE, size = 4) +
   scale_color_manual(values = c("goldenrod", "forestgreen",  "royalblue", "darkorchid")) +
[…]

### 2022-01-05 src/figure-spp-changes.R

@@ -1,3 +1,5 @@
+library(tidyverse)
+
 ### NMDS species scores
 
 read.csv("data/spatial-survey-species.csv") %>%
@@ -17,6 +19,29 @@ read.csv("data/spatial-survey-species.csv") %>%
   mutate(Species = paste(substr(G, 1, 3), substr(S, 1, 3), sep = "")) %>%
   dplyr::select(Taxon, Species, Dim.1, Dim.2) -> sppscores
 
+### SNCs
+
+read.csv("data/spatial-survey-temperatures.csv") %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
+  group_by(Plot, Day = lubridate::floor_date(Time, "day")) %>%
+  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
+  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
+  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
+  group_by(Plot, Month = lubridate::floor_date(Day, "month")) %>%
+  summarise(FDD = sum(FDD), # FDD per month
+            GDD = sum(GDD)) %>% # GDD per month
+  group_by(Plot) %>%
+  summarise(FDD = sum(FDD), # FDD per year
+            GDD = sum(GDD))  %>%
+  select(Plot, FDD, GDD) %>%
+  merge(read.csv("data/spatial-survey-species.csv")) %>%
+  group_by(Taxon) %>%
+  summarise(n = length(Plot),
+            GDD = weighted.mean(GDD, Cover),
+            FDD = abs(weighted.mean(FDD, Cover))) %>%
+  # filter(n > 4) %>%
+  arrange(Taxon) -> sncs
+
 ### Species changes
 
 read.csv("data/temporal-survey-matrices.csv") %>%
@@ -33,50 +58,33 @@ read.csv("data/temporal-survey-matrices.csv") %>%
   mutate(ChangeP = `19` / `09`) %>%
   arrange(Change) %>%
   na.omit %>%
-  merge(sppscores) -> changespp
+  merge(sppscores) %>%
+  merge(sncs) %>%
+  mutate(C = ifelse(Change < -5, "Decreasing frequency", "Minimal changes"),
[…]

### 2022-01-05 src/figure-survey-sites.R

@@ -85,66 +85,19 @@ tPlots %>%
 
 ### Panel C - The Crosses
 
-img1 <- png::readPNG("map/survey-design.png")
+img1 <- png::readPNG("map/survey-design2.png")
 
-f3 <- grid::rasterGrob(img1, width = unit(120, "mm"), height = unit(43, "mm"), just = "centre")
-
-### Panel D - The Climographs
-
-read.csv("data/temporal-survey-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
-  as_tbl_time(index = Time) %>%
-  arrange(Time) %>%
-  filter_time("2009" ~ "2018") %>%
-  mutate(m = month(Time)) %>%
-  group_by(Site, m) %>%
-  summarise(T = mean(Temperature)) -> ts # Monthly average temperature
-
-read.csv("data/temporal-survey-wp.csv") %>%
-  mutate(m = month(Time)) %>%
-  group_by(Site, m) %>%
-  summarise(WP = mean(WP)) -> wp # Monthly average water potential
-
-merge(ts, wp) %>%
-  mutate(Site = fct_relevel(Site,
-                            "Los Cazadores", "Hou Sin Tierri",
-                            "Los Boches", "Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) %>%
-  ggplot(aes(x = m)) +
-  facet_wrap(~ Site, nrow = 1) +
-  geom_bar(aes(y = -WP*10), stat = "identity", position = "dodge", fill = "dodgerblue4", size = 1.1) +
-  geom_line(aes(y = T), color = "darkred", size = .5) +
-  geom_point(aes(y = T), color = "darkred", size = .7) +
-  scale_y_continuous(sec.axis = sec_axis( trans = ~./10, name = "Mean monthly soil water potential (-MPa)")) +
-  ggthemes::theme_tufte() +
-  theme(plot.title = element_text(hjust = -.15, vjust = -1.5),
-        axis.ticks.y = element_line(colour = "darkred"),
-        axis.ticks.y.right = element_line(colour = "dodgerblue4"),
-        axis.text.y = element_text(colour = "darkred"),
-        axis.text.y.right = element_text(colour = "dodgerblue4"),
-        axis.title.y = element_text(colour = "darkred"),
-        axis.title.y.right = element_text(colour = "dodgerblue4"),
[…]

### 2022-01-05 src/figure-temperature-changes.R

@@ -1,57 +1,0 @@
-library(tidyverse); library(urca)
-
-### Decompose time series
-
-decomposer <- function(x) 
-{
-  x %>% 
-    pull(Temperature) %>%
-    ts(start = c(2009, 1), frequency = 24 * 365) -> timeseries
-  
-  stl(timeseries, s.window = "period") -> fit
-  
-  cbind(x$Time, as.data.frame(timeseries), as.data.frame(fit$time.series)) -> df
-  
-  colnames(df) <- c("Time", "Temperature", "Seasonal", "Trend", "Remainder")
-  
-  df
-}
-
-read.csv("data/temporal-survey-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
-  group_by(Site) %>%
-  do(decomposer(.)) %>%
-  mutate(Site = as.factor(Site)) %>%
-  mutate(Site = fct_relevel(Site,
-                            "Los Cazadores", "Hou Sin Tierri",
-                            "Los Boches", "Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> trends
-
-### Plot trends
-
-trends %>%
-  ggplot(aes(Time, Trend)) + 
-  facet_wrap(~ Site, nrow = 1) + 
-  geom_line(color = "dodgerblue") +
-  geom_smooth(method = "lm", color = "dodgerblue4") +
-  xlab("Time (1-h recording inverval)") +
-  ylab("Trend (ºC)") +
-  ggthemes::theme_tufte() +
-  theme(legend.position = "none", 
-        legend.direction = "vertical",
-        legend.title = element_blank(),
-        legend.text = element_text(size = 16, face = "italic"), 
-        panel.background = element_rect(color = "grey96", fill = "grey96"),
-        strip.text = element_text(size = 12),
-        axis.title = element_text(size = 12),
-        axis.text.x = element_text(size = 8, color = "black"),
-        axis.text.y = element_text(size = 12, color = "black")) -> f1; f1
-
-### Save figure
-
-ggsave(f1, file = "results/figures/temperature-changes.png", 
-       path = NULL, scale = 1, width = 182, height = 182/3, units = "mm", dpi = 600)
-# ggsave(f1, file = "results/figures/temperature-changes.tiff", device = grDevices::tiff, 
[…]

### 2022-01-05 src/figure-temperature-deviations.R

@@ -1,0 +1,90 @@
+library(tidyverse); library(lubridate)
+
+### Temporal survey data
+
+read.csv("data/temporal-survey-temperatures.csv") %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
+  group_by(Site, Day = lubridate::floor_date(Time, "day")) %>%
+  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>%
+  mutate(d = day(Day), 
+         m = month(Day)) -> years
+
+years %>%
+  group_by(Site, m, d) %>%
+  summarise(TM = mean(T)) %>%
+  merge(years) %>% 
+  mutate(Site = fct_relevel(Site,
+                            "Los Cazadores", "Hou Sin Tierri",
+                            "Los Boches", "Hoyo Sin Tierra")) %>%
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> time
+
+### Spatial survey data
+
+read.csv("data/spatial-survey-temperatures.csv") %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
+  merge(read.csv("data/spatial-survey-header.csv")) %>%
+  group_by(Plot, Site, Day = lubridate::floor_date(Time, "day")) %>%
+  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) -> sites
+
+sites %>%
+  group_by(Site, Day) %>%
+  summarise(TM = mean(T)) %>%
+  merge(sites) %>% 
+  mutate(Site = fct_relevel(Site,
+                            "Los Cazadores", "Hou Sin Tierri",
+                            "Los Boches", "Hoyo Sin Tierra")) %>%
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> space
+
+### Temporal survey deviations
+
+time %>%
+  ggplot(aes(TM, T, fill = Site)) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  geom_vline(xintercept = 0, linetype = "dashed") +
+  geom_point(alpha = 0.5, shape = 21) +
+  geom_abline(intercept = 0, slope = 1, size = 1) +
+  scale_fill_manual(values = c("goldenrod", "forestgreen",  "royalblue", "darkorchid")) +
+  ggthemes::theme_tufte() + 
[…]

### 2022-01-05 src/figure-temperature-logs.R

@@ -1,4 +1,4 @@
-library(tidyverse)
+library(tidyverse); library(urca); library(raster); library(tibbletime); library(lubridate)
 Sys.setlocale("LC_TIME", "English")
 
 ### Geoprecision logs
@@ -12,12 +12,13 @@ read.csv("data/temporal-survey-temperatures.csv") %>%
   ggplot(aes(Time, Temperature)) + 
   facet_wrap(~ Site, nrow = 1) + 
   geom_line(color = "red") +
-  labs(title = "(A) Temperature logs from the temporal survey (2009-2018)") +
+  labs(title = "(C) Hourly logs, temporal survey (2009-2018)") +
   xlab("Time (1-h recording inverval)") +
   ylab("Temperature (ºC)") +
   geom_hline(yintercept = 0, linetype = "dashed") +
   ggthemes::theme_tufte() +
-  theme(legend.position = "none", 
+  theme(strip.background = element_rect(colour = "grey96", fill = "grey96"),
+        legend.position = "none", 
         legend.direction = "vertical",
         legend.title = element_blank(),
         legend.text = element_text(size = 16, face = "italic"), 
@@ -25,7 +26,7 @@ read.csv("data/temporal-survey-temperatures.csv") %>%
         strip.text = element_text(size = 12),
         axis.title = element_text(size = 12),
         axis.text.x = element_text(size = 8, color = "black"),
-        axis.text.y = element_text(size = 12, color = "black")) -> f1; f1
+        axis.text.y = element_text(size = 11, color = "black")) -> f1; f1
 
 ### iButton logs
 
@@ -40,26 +41,121 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
   facet_wrap(~ Site, nrow = 1) + 
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_line() +
-  labs(title = "(B) Temperature logs from the spatial survey (20 iButtons per site)") +
+  labs(title = "(D) Hourly logs, spatial survey (20 iButtons per site, Oct 2018 - Aug 2019)") +
   xlab("Time (4-h recording inverval)") +
   ylab("Temperature (ºC)") +
   ggthemes::theme_tufte() +
-  theme(legend.position = "none", 
+  theme(strip.background = element_rect(colour = "grey96", fill = "grey96"),
+        legend.position = "none", 
         legend.direction = "vertical",
[…]

### 2022-01-05 src/figures-spacevstime.R

@@ -1,0 +1,168 @@
+library(tidyverse); library(tibbletime); library(vegan)
+
+### Indices for GeoPrecision by complete year
+
+read.csv("data/temporal-survey-temperatures.csv") %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
+  group_by(Site, Day = lubridate::floor_date(Time, "day")) %>%
+  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
+  mutate(Snow = ifelse(X < 0.5 & N > -0.5, 1, 0)) %>% # Day is snow day or not
+  # mutate(Snow = ifelse((X - N) <= 0.5, 1, 0)) %>% # Day is snow day or not
+  mutate(FreezeThaw = ifelse(X > 0.5 & N < -0.5, 1, 0)) %>% # Day with freeze-thaw cycles
+  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
+  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
+  group_by(Site, Month = lubridate::floor_date(Day, "month")) %>%
+  summarise(T = mean(T), X = mean(X), N = mean(N), # Daily mean, max, min
+            Snow = sum(Snow), # Snow days per month
+            FreezeThaw = sum(FreezeThaw), # Freeze-thaw days per month
+            FDD = sum(FDD), # FDD per month
+            GDD = sum(GDD)) %>% # GDD per month
+  group_by(Site, Year = lubridate::floor_date(Month, "year")) %>%
+  summarise(bio01 = mean(T), # Annual Mean Temperature
+            bio02 = mean(X - N), # Mean Diurnal Range (Mean of monthly (max temp - min temp))
+            bio04 = sd(T) * 100, # Temperature Seasonality (standard deviation ×100)
+            bio05 = max(X), # Max Temperature of Warmest Month
+            bio06 = min(N), # Min Temperature of Coldest Month
+            bio07 = bio05 - bio06, # Temperature Annual Range (BIO5-BIO6)
+            bio03 = (bio02 / bio07) * 100, # Isothermality (BIO2/BIO7) (×100)
+            Snow = sum(Snow), # Days of snow per year
[…]

### 2022-01-05 src/indices.R

@@ -1,134 +1,0 @@
-library(tidyverse); library(tibbletime)
-
-merge(read.csv("data/logs.csv"),
-      read.csv("data/loggers.csv")) %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
-  as_tbl_time(index = Time) %>%
-  arrange(Brand, Site, Time) -> logs
-
-# End and start dates
-
-logs %>%
-  filter(Brand == "GeoPrecision") %>%
-  group_by(Site) %>%
-  summarise(Min = min(Time), Max = max(Time))
-
-logs %>%
-  filter(Brand == "EMS") %>%
-  group_by(Site) %>%
-  summarise(Min = min(Time), Max = max(Time))
-
-logs %>%
-  filter(Brand == "iButtons") %>%
-  group_by(Site) %>%
-  summarise(Min = min(Time), Max = max(Time))
-
-# Compare temperature GeoPrecision vs. EMS
-
-logs %>%
-  filter(Brand %in% c("GeoPrecision", "EMS")) %>%
-  filter(Site == c("Urriellu")) %>%
-  filter(Time > "2017-01-01") %>%
-  ggplot(aes(Time, Temperature, color = Brand)) +
-  geom_line() + 
-  facet_wrap(~ Site) # Huge differences in Hoyo Sin Tierra
-
-# Indices for GeoPrecision by complete year
-
-logs %>%
-  filter(Brand == "GeoPrecision" | (Brand == "EMS" & Site == "Urriellu")) %>% # Use EMS to fill in missing GeoPrecision data at Urriellu
-  arrange(Time) %>%
-  filter_time("2009" ~ "2018") %>%
-  group_by(Site, Day = lubridate::floor_date(Time, "day")) %>%
-  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
-  mutate(Snow = ifelse(X < 0.5 & N > -0.5, 1, 0)) %>% # Day is snow day or not
-  # mutate(Snow = ifelse((X - N) <= 0.5, 1, 0)) %>% # Day is snow day or not
-  mutate(FreezeThaw = ifelse(X > 0.5 & N < -0.5, 1, 0)) %>% # Day with freeze-thaw cycles
-  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
[…]

### 2022-01-05 src/sncs.R

@@ -1,48 +1,0 @@
-library(tidyverse); library(tibbletime)
-
-read.csv("data/focal-plot-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) -> temperatures
-
-read.csv("data/ibutton-plot-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
-  merge(read.csv("data/ibutton-plot-header.csv")) %>%
-  group_by(Plot, Site, Day = lubridate::floor_date(Time, "day")) %>%
-  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
-  mutate(Snow = ifelse(X < 0.5 & N > -0.5, 1, 0)) %>% # Day is snow day or not
-  mutate(FreezeThaw = ifelse(X > 0.5 & N < -0.5, 1, 0)) %>% # Day with freeze-thaw cycles
-  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
-  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
-  group_by(Plot, Site, Month = lubridate::floor_date(Day, "month")) %>%
-  summarise(T = mean(T), X = mean(X), N = mean(N), # Daily mean, max, min
-            Snow = sum(Snow), # Snow days per month
-            FreezeThaw = sum(FreezeThaw), # Freeze-thaw days per month
-            FDD = sum(FDD), # FDD per month
-            GDD = sum(GDD)) %>% # GDD per month
-  group_by(Plot, Site) %>%
-  summarise(bio01 = mean(T), # Annual Mean Temperature
-            bio02 = mean(X - N), # Mean Diurnal Range (Mean of monthly (max temp - min temp))
-            bio04 = sd(T) * 100, # Temperature Seasonality (standard deviation ×100)
-            bio05 = max(X), # Max Temperature of Warmest Month
-            bio06 = min(N), # Min Temperature of Coldest Month
-            bio07 = bio05 - bio06, # Temperature Annual Range (BIO5-BIO6)
-            bio03 = (bio02 / bio07) * 100, # Isothermality (BIO2/BIO7) (×100)
-            Snow = sum(Snow), # Days of snow per year
[…]

### 2022-01-05 src/sorensen.R

@@ -1,53 +1,0 @@
-library(tidyverse); library(vegan)
-  
-sor <- function(x)  
-  {x %>%
-  select(Plot, Taxon, Cover) %>%
-  spread(Taxon, Cover, fill = 0) %>%
-  column_to_rownames(var = "Plot") %>%
-  vegdist(method = "bray", binary = TRUE) %>%
-  mean %>%
-    data.frame}
-
-read.csv("data/ibutton-plot-species.csv") %>%
-  merge(read.csv("data/ibutton-plot-header.csv")) %>%
-  group_by(Site) %>%
-  do(sor(.)) %>%
-  rename(Space = ".") -> inspace
-
-read.csv("data/focal-plot-species.csv") %>%
-  merge(read.csv("data/focal-plot-header.csv")) %>%
-  separate(Plot, into = c("Plot2", "Year"), sep = "-", remove = FALSE) %>%
-  group_by(Site, Plot2) %>%
-  do(sor(.)) %>%
-  rename(Time = ".") %>%
-  group_by(Site) %>%
-  summarise(Time = mean(Time)) -> intime
-
-merge(intime, inspace) %>%
-  mutate(Site = fct_relevel(Site, "El Cable", "Hoyo Sin Tierra", "Cabaña Verónica",
-                            "Horcados Rojos", "Los Boches", "Hou Sin Tierri",
-                            "Los Cazadores", "Urriellu")) %>%
-  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) %>%
-  gather(Dimension, Value, Time:Space) %>%
-  ggplot(aes(Dimension, Value, fill = Dimension)) + 
-  facet_wrap(~ Site, nrow = 1) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Species dissimilarity: in space (iButtons) vs. time (10 years)") +
-  xlab("Sites") +
-  ylab("Sørensen dissimilarity index (1 - Sørensen)") +
-  ggthemes::theme_tufte() +
-  scale_fill_manual(values = c("purple", "gold")) + 
-  theme(legend.position = "right", 
-        legend.title = element_blank(),
-        legend.text = element_text(size = 16, face = "italic"), 
-        strip.text = element_text(size = 22), 
-        panel.background = element_rect(color = "grey96", fill = "grey96"),
-        title = element_text(size = 24, color = "black", face = "bold"),
-        axis.title.y = element_text(size = 22, color = "black"),
[…]

### 2022-01-05 src/temperature-differences.R

@@ -1,188 +1,0 @@
-library(tidyverse); library(tibbletime)
-
-read.csv("data/focal-plot-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) -> temperatures
-
-# Indices for GeoPrecision by complete year
-
-temperatures %>%
-  group_by(Site, Day = lubridate::floor_date(Time, "day")) %>%
-  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
-  mutate(Snow = ifelse(X < 0.5 & N > -0.5, 1, 0)) %>% # Day is snow day or not
-  # mutate(Snow = ifelse((X - N) <= 0.5, 1, 0)) %>% # Day is snow day or not
-  mutate(FreezeThaw = ifelse(X > 0.5 & N < -0.5, 1, 0)) %>% # Day with freeze-thaw cycles
-  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
-  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
-  group_by(Site, Month = lubridate::floor_date(Day, "month")) %>%
-  summarise(T = mean(T), X = mean(X), N = mean(N), # Daily mean, max, min
-            Snow = sum(Snow), # Snow days per month
-            FreezeThaw = sum(FreezeThaw), # Freeze-thaw days per month
-            FDD = sum(FDD), # FDD per month
-            GDD = sum(GDD)) %>% # GDD per month
-  group_by(Site, Year = lubridate::floor_date(Month, "year")) %>%
-  summarise(bio01 = mean(T), # Annual Mean Temperature
-            bio02 = mean(X - N), # Mean Diurnal Range (Mean of monthly (max temp - min temp))
-            bio04 = sd(T) * 100, # Temperature Seasonality (standard deviation ×100)
-            bio05 = max(X), # Max Temperature of Warmest Month
-            bio06 = min(N), # Min Temperature of Coldest Month
-            bio07 = bio05 - bio06, # Temperature Annual Range (BIO5-BIO6)
-            bio03 = (bio02 / bio07) * 100, # Isothermality (BIO2/BIO7) (×100)
-            Snow = sum(Snow), # Days of snow per year
[…]

### 2022-01-05 src/vegetation-changes.R

@@ -1,241 +1,0 @@
-library(tidyverse)
-
-read.csv("data/focal-plot-matrices.csv") -> matrices
-
-matrices %>%
-  merge(read.csv("data/focal-plot-header.csv")) %>%
-  filter(Site %in% c("Hou Sin Tierri", "Hoyo Sin Tierra", "Los Boches", "Los Cazadores")) %>%
-  group_by(Plot, Taxon) %>%
-  summarise(Cover = sum(Present)) %>%
-  merge(read.csv("data/checklist.csv"), all.x = TRUE) %>%
-  filter(Lifeform != "Therophyte") %>%
-  separate(Plot, into = c("Plot", "Year"), sep = "-") %>%
-  mutate(Year = fct_recode(Year, "09" = "08")) %>%
-  mutate(Taxon = fct_recode(Taxon,
-                            "Ranunculus favargeri" = "Ranunculus parnassiifolius subsp. favargeri",
-                            "Dethawia cantabrica" = "Dethawia splendens subsp. cantabrica",
-                            "Helianthemum urrielense" = "Helianthemum apenninum subsp. urrielense",
-                            "Thymus ligusticus" = "Thymus praecox subsp. ligusticus")) %>%
-  group_by(Taxon, Year) %>%
-  summarise(Cover = sum(Cover)) %>%
-  spread(Year, Cover) %>%
-  mutate(Change = `19` - `09`) %>%
-  arrange(Change) %>%
-  filter(Change < -10 | Change > 10) %>%
-  na.omit %>%
-  ggplot(aes(reorder(Taxon, Change), Change)) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  geom_bar(stat = "identity", fill = "limegreen") +
-  labs(title = "10-year species changes") +
-  xlab("Species") +
-  ylab("Change (in 10x10 cm presences)") +
-  ggthemes::theme_tufte() +
-  theme(legend.position = "none", 
-        legend.direction = "vertical",
-        legend.title = element_blank(),
-        legend.text = element_text(size = 16, face = "italic"), 
-        strip.text = element_text(size = 22), 
-        panel.background = element_rect(color = "grey96", fill = "grey96"),
-        title = element_text(size = 24, color = "black", face = "bold"),
-        axis.title.y = element_text(size = 22, color = "black"),
[…]

### 2022-01-14 doc/manuscript.Rmd
1st manuscript draft 🌅😊

---
title: "Microclimatic spatial variation buffers local warming in alpine vegetation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

B Jiménez-Alfaro, E Fernández-Pascual, A Bueno, C Marcenó….¿?

# Abstract (ME&B)

Alpine vegetation is considered to respond to climate change through species distribution shifts with effects on community richness and composition. However, the complex topography of alpine landscapes creates a mosaic of microclimatic niches which might buffer macroclimatic variation, preventing local extinctions of alpine species. The magnitude of microclimatic buffering is context-dependent and may depend on regional factors such as biogeographical settings, topography, or species pools. A key question for understanding the magnitude of this buffering is therefore to test whether spatial topographic variation reflects temporal changes within and between years in different regions. We addressed this question in a long-term vegetation monitoring site in Picos de Europa National Park, a transitional mountain massif between the temperate and Mediterranean macroclimates in northern Spain. In 2008, we established four stations along an alpine landscape for measuring temporal changes in soil temperatures and related changes in plant communities based on two 1 m2 plots replicated around the central point. In 2018, we also measured spatial variation in soil temperatures within each station using 20 additional plots placed at 10 m intervals along with cardinal directions from the central point. We found that spatial variation was higher than 10-year temporal variation for most of the bioclimatic indices we used, but there were exceptions in stations that accumulate less snow, and in less topographically diverse stations. Vegetation variation (Sørensen index) was higher across topographical gradients than across time. These results suggest that microclimatic refugia can compensate for temporal changes, but this compensation might not be homogeneous throughout the alpine landscape. The response of the study system to contemporary climate warming seems more likely to produce a slow re-accommodation of species relative abundances along with topographical variation, rather than local extinctions.

# Introduction

xxx

# Methods

## Study system

The study was conducted in the calcareous massif of Picos de Europa National Park, in northern Spain (Figure 1A). The study area is a biodiversity hotspot for cold-adapted plants in the Iberian Peninsula and a biogeographical hub for Alpine and Mediterranean lineages in Western Europe (XXX). The main calcareous massif occupies c. 50 km2 and support high diversity of ecosystems, with alpine vegetation mostly occurring between 1900 and 2400 m a.s.l., with a local species pool of XX species (Jiménez-Alfaro et al. XXX). In 2008, we established a long-term ecological research program for monitoring soil climate and vegetation change. We selected four study sites along a north-south gradient, reflecting the main macroclimatic variation from oceanic to Mediterranean influence. In 2016, we additionally installed a Water potential logger (XXX) to characterize water stress in the four sites and to ….  The sites were selected to… general design and description of sites… (Borja)

> Empiezo a pensar que sería mejor quitar esto, por mucho que me fastidie, y guardarnos el tema del WP para otro paper junto con lo de Omaña. Eso simplificaría el gráfico (de otro modo, tendría más lógica ponerlo con la Figura 1, porque en realidad no hacemos nada con esto)

## Vegetation and microclimatic data

In each site, we buried a temperature logger (XXXX) at 5 cm depth  in a relatively flat and homogeneous vegetation patch. We surveyed the plant community in two replicated plots of 1 m2 separated 1 m from the logger, identifying species composition and estimating relative cover in %. In each plot, we installed a grid template of 100 microplots (10 x 10 cm each) to sample species frequency according to the standard methodology of GLORIA (XXX). The loggers were replaced by new ones, when needed, to obtain a continuous temperature record from 2008 to 2018. In 2018, we resampled the same plots in the same way to detect potential changes in species presence and frequency. The vegetation data from these surveys, together with the soil temperature collected in the four study sites during 10 years, represent the “temporal survey”. In 2019, we visited the same areas to study the spatial variation of vegetation and microclimate around the previously sampled areas. Using the long-term temperature logger as the central point, we additionally placed 20 ibuttons (XXX) in 20 plots of 1 m2 separated 10 m from each other in the four cardinal directions (Figure 1C).. etc… (Borja)

## Data analysis

Vegetation data + Temperature data (Eduardo)

# Results

## Microclimatic variation

Along the 10 years of soil temperature monitoring, we found a general trend of increasing temperature, especially in the southwest and warmest site (Hoyo sin Tierra) (Figure 3). Two of the sites (Los Cazadores and Los Boches) showed a consistent pattern of continuous snow cover during winter (reflected by temperature records around 0º C). In contrast, the two other sites (Hoy Sin Tierri and Hoyo Sin Tierra) showed frost temperatures during most winters (although they may not be constant and intermixed with short snow periods). Maximum temperatures were… Soil temperature corresponding to the spatial surveys also showed high variation within sites (Figure 3). In the four sites we found plots with contrasting patterns of snow cover and frost during winter, reflected by…  When looking at the whole microclimatic variation among plots (Figure 4), we found that the four sites were represented by a wide range of temperatures, with the exception of one site (Los Cazadores). The two main axes of variation were mainly related to FDD and GDD, the latter also correlated with…. (Eduardo)

## Plant diversity and microclimate

In the temporal survey (2 plots per site, n =16), we recorded XX species in 2008 and XX in 2018. (species loss, gains)… Changes in species frequency were/were not correlated with XX and XX…. (Table 1) (Eduardo). In the spatial survey (20 plots per site, n = 80; including four plots placed in rocks with no plants) we recorded XX species. Changes in species composition were associated to GDD (NMDS axis 1) to represent a main gradient from the warmest to the coldest sites (Figure 4), and to FDD (NMDS2) in a gradient of snow versus frost winters. In the compositional space, the sites were poorly differentiated across the four sites (PERMANOVA, adonis) mainly by the effect of one site (Los Boches) with most plots places in the cold and frost margins. Etc…. (Borja).

## Microclimatic ranges in time and space

…. (Eduardo) (quizás repensar esta parte, pero es importante mantenerla porque es la clave).

### 2022-01-14 doc/manuscript.Rmd
changes to ms

@@
 title: "Microclimatic spatial variation buffers local warming in alpine vegetation"
-output: html_document
+output: github_document
+bibliography: refs.bib
+csl: annals-of-botany.csl
+link-citations: yes
+always_allow_html: true
 ---
@@
 
-B Jiménez-Alfaro, E Fernández-Pascual, A Bueno, C Marcenó….¿?
+B Jiménez-Alfaro, E Fernández-Pascual, A Bueno, C Marcenó....¿?
 
@@
 We selected four study sites along a north-south gradient, reflecting the main macroclimatic variation from oceanic to Mediterranean influence.
-In 2016, we additionally installed a Water potential logger (XXX) to characterize water stress in the four sites and to ….
-The sites were selected to… general design and description of sites… (Borja)
+In 2016, we additionally installed a Water potential logger (XXX) to characterize water stress in the four sites and to ....
+The sites were selected to...
+general design and description of sites...
+(Borja)
 
@@
 In 2018, we resampled the same plots in the same way to detect potential changes in species presence and frequency.
-The vegetation data from these surveys, together with the soil temperature collected in the four study sites during 10 years, represent the “temporal survey”.
+The vegetation data from these surveys, together with the soil temperature collected in the four study sites during 10 years, represent the "temporal survey".
 In 2019, we visited the same areas to study the spatial variation of vegetation and microclimate around the previously sampled areas.
 Using the long-term temperature logger as the central point, we additionally placed 20 ibuttons (XXX) in 20 plots of 1 m2 separated 10 m from each other in the four cardinal directions (Figure 1C)..
-etc… (Borja)
+etc...
+(Borja)
 
 ## Data analysis
+
+[@RN2315]
 
@@
 In contrast, the two other sites (Hoy Sin Tierri and Hoyo Sin Tierra) showed frost temperatures during most winters (although they may not be constant and intermixed with short snow periods).
-Maximum temperatures were… Soil temperature corresponding to the spatial surveys also showed high variation within sites (Figure 3).
-In the four sites we found plots with contrasting patterns of snow cover and frost during winter, reflected by…  When looking at the whole microclimatic variation among plots (Figure 4), we found that the four sites were represented by a wide range of temperatures, with the exception of one site (Los Cazadores).
-The two main axes of variation were mainly related to FDD and GDD, the latter also correlated with….
+Maximum temperatures were...
+Soil temperature corresponding to the spatial surveys also showed high variation within sites (Figure 3).
+In the four sites we found plots with contrasting patterns of snow cover and frost during winter, reflected by...
+When looking at the whole microclimatic variation among plots (Figure 4), we found that the four sites were represented by a wide range of temperatures, with the exception of one site (Los Cazadores).
+The two main axes of variation were mainly related to FDD and GDD, the latter also correlated with....
 (Eduardo)
@@
 In the temporal survey (2 plots per site, n =16), we recorded XX species in 2008 and XX in 2018.
-(species loss, gains)… Changes in species frequency were/were not correlated with XX and XX….
+(species loss, gains)...
+Changes in species frequency were/were not correlated with XX and XX....
 (Table 1) (Eduardo).
@@
 In the compositional space, the sites were poorly differentiated across the four sites (PERMANOVA, adonis) mainly by the effect of one site (Los Boches) with most plots places in the cold and frost margins.
-Etc….
+Etc....
 (Borja).
@@
 
-…. (Eduardo) (quizás repensar esta parte, pero es importante mantenerla porque es la clave).
+.... (Eduardo) (quizás repensar esta parte, pero es importante mantenerla porque es la clave).

### 2022-01-19 src/figure-pca-temperatures.R
4 x PCA

@@ -1,9 +1,45 @@
 library(tidyverse)
 
+### Temporal survey
+
+read.csv("data/temporal-survey-temperatures.csv") %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
+  mutate(Month = lubridate::month(Time)) %>% 
+  mutate(Day = lubridate::day(Time)) %>% 
+  mutate(Hour = lubridate::hour(Time)) %>% 
+  filter(Hour %in% c(0, 4, 8, 12, 16, 20)) %>% # Keep same recording hours as ibuttons
+  filter(! Month %in% 9) %>% # Remove missing September days
+  filter(! (Month %in% 8 & Day > 6)) %>% # Remove missing August days
+  filter(! (Month %in% 10 & Day < 3)) %>% # Remove missing October days %>%
+  dplyr::select(-c(Month, Day, Hour)) %>%
+  group_by(Site, Day = lubridate::floor_date(Time, "day")) %>%
+  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
+  mutate(Snow = ifelse(X < 0.5 & N > -0.5, 1, 0)) %>% # Day is snow day or not
+  # mutate(Snow = ifelse((X - N) <= 0.5, 1, 0)) %>% # Day is snow day or not
+  mutate(FreezeThaw = ifelse(X > 0.5 & N < -0.5, 1, 0)) %>% # Day with freeze-thaw cycles
+  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
+  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
+  group_by(Site, Month = lubridate::floor_date(Day, "month")) %>%
+  summarise(T = mean(T), X = mean(X), N = mean(N), # Daily mean, max, min
+            Snow = sum(Snow), # Snow days per month
+            FreezeThaw = sum(FreezeThaw), # Freeze-thaw days per month
+            FDD = sum(FDD), # FDD per month
+            GDD = sum(GDD)) %>% # GDD per month
+  group_by(Site, Year = lubridate::floor_date(Month, "year")) %>%
+  summarise(bio1 = mean(T), # Annual Mean Temperature
[…]

### 2022-01-19 src/figure-temperature-logs.R

@@ -1,33 +1,6 @@
 library(tidyverse); library(urca); library(raster); library(tibbletime); library(lubridate)
 Sys.setlocale("LC_TIME", "English")
 
-### Geoprecision logs
-
-read.csv("data/temporal-survey-temperatures.csv") %>% 
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
-  mutate(Site = fct_relevel(Site,
-                            "Los Cazadores", "Hou Sin Tierri",
-                            "Los Boches", "Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) %>%
-  ggplot(aes(Time, Temperature)) + 
-  facet_wrap(~ Site, nrow = 1) + 
-  geom_line(color = "red") +
-  labs(title = "(C) Hourly logs, temporal survey (2009-2018)") +
-  xlab("Time (1-h recording inverval)") +
-  ylab("Temperature (ºC)") +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  ggthemes::theme_tufte() +
-  theme(strip.background = element_rect(colour = "grey96", fill = "grey96"),
-        legend.position = "none", 
-        legend.direction = "vertical",
-        legend.title = element_blank(),
-        legend.text = element_text(size = 16, face = "italic"), 
-        panel.background = element_rect(color = "grey96", fill = "grey96"),
-        strip.text = element_text(size = 12),
-        axis.title = element_text(size = 12),
-        axis.text.x = element_text(size = 8, color = "black"),
-        axis.text.y = element_text(size = 11, color = "black")) -> f1; f1
-
 ### iButton logs
 
 read.csv("data/spatial-survey-temperatures.csv") %>%
@@ -41,10 +14,39 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
   facet_wrap(~ Site, nrow = 1) + 
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_line() +
-  labs(title = "(D) Hourly logs, spatial survey (20 iButtons per site, Oct 2018 - Aug 2019)") +
+  labs(title = "(A) Hourly logs, spatial survey (20 iButtons per site, Oct 2018 - Aug 2019)") +
   xlab("Time (4-h recording inverval)") +
   ylab("Temperature (ºC)") +
   ggthemes::theme_tufte() +
[…]

### 2022-01-19 src/figure-nmds-species.R
new nmds

@@ -4,10 +4,27 @@ library(tidyverse)
 
 read.csv("data/spatial-survey-species.csv") %>%
   filter(Plot != "A03") %>%
-  filter(Plot != "D19") %>%
+  filter(Plot != "D19") -> speciesSS # Vegetation data from the spatial survey
+
+read.csv("data/temporal-survey-species.csv")  -> speciesTS # Vegetation data from the temporal survey
+ 
+rbind(speciesSS, speciesTS)  %>%
   spread(Taxon, Cover, fill = 0) %>%
   arrange(Plot) %>%
-  column_to_rownames(var = "Plot") -> species
+  column_to_rownames(var = "Plot") -> species # Merge species
+
+### Header data
+
+read.csv("data/spatial-survey-header.csv") %>% mutate(Survey = "Spatial") %>%
+  merge(read.csv("results/numerical/indices.csv")) -> headerSS # Header data from the spatial survey
+
+read.csv("results/numerical/indices.csv") %>%
+  filter(Survey == "Temporal") %>%
+  filter(Plot == "2018") %>% 
+  select(-Plot) %>% 
+  merge(read.csv("data/temporal-survey-header.csv")) -> headerTS # Header data from the temporal survey
+
+rbind(headerSS, headerTS) -> header # Merge header
 
 ### Do NMDS
 
@@ -21,8 +38,8 @@ vegan::scores(nmds, "sites") %>%
   data.frame() %>%
   rownames_to_column("Plot") %>%
   rename(Dim.1 = NMDS1, Dim.2 = NMDS2) %>%
-  merge(read.csv("data/spatial-survey-header.csv")) %>%
-  dplyr::select(Plot, Site, Dim.1, Dim.2) %>% 
+  merge(header) %>%
+  dplyr::select(Plot, Site, Survey, Dim.1, Dim.2) %>% 
   mutate(Site = fct_relevel(Site,
                             "Los Cazadores", "Hou Sin Tierri",
                             "Los Boches", "Hoyo Sin Tierra")) %>%
@@ -30,16 +47,9 @@ vegan::scores(nmds, "sites") %>%
 
 ### Species scores
 
-read.csv("data/spatial-survey-species.csv") %>%
-  group_by(Taxon) %>%
-  tally %>%
-  filter(n > 80 * .2) %>%
-  pull(Taxon) -> sppfreq
-
 vegan::scores(nmds, "species") %>%
   data.frame() %>%
   rownames_to_column("Taxon") %>%
[…]

### 2022-01-19 src/figure-pca-temperatures.R

@@ -1,70 +1,6 @@
 library(tidyverse)
 
-### Temporal survey
-
-read.csv("data/temporal-survey-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
-  mutate(Month = lubridate::month(Time)) %>% 
-  mutate(Day = lubridate::day(Time)) %>% 
-  mutate(Hour = lubridate::hour(Time)) %>% 
-  filter(Hour %in% c(0, 4, 8, 12, 16, 20)) %>% # Keep same recording hours as ibuttons
-  filter(! Month %in% 9) %>% # Remove missing September days
-  filter(! (Month %in% 8 & Day > 6)) %>% # Remove missing August days
-  filter(! (Month %in% 10 & Day < 3)) %>% # Remove missing October days %>%
-  dplyr::select(-c(Month, Day, Hour)) %>%
-  group_by(Site, Day = lubridate::floor_date(Time, "day")) %>%
-  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
-  mutate(Snow = ifelse(X < 0.5 & N > -0.5, 1, 0)) %>% # Day is snow day or not
-  # mutate(Snow = ifelse((X - N) <= 0.5, 1, 0)) %>% # Day is snow day or not
-  mutate(FreezeThaw = ifelse(X > 0.5 & N < -0.5, 1, 0)) %>% # Day with freeze-thaw cycles
-  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
-  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
-  group_by(Site, Month = lubridate::floor_date(Day, "month")) %>%
-  summarise(T = mean(T), X = mean(X), N = mean(N), # Daily mean, max, min
-            Snow = sum(Snow), # Snow days per month
-            FreezeThaw = sum(FreezeThaw), # Freeze-thaw days per month
-            FDD = sum(FDD), # FDD per month
-            GDD = sum(GDD)) %>% # GDD per month
-  group_by(Site, Year = lubridate::floor_date(Month, "year")) %>%
[…]

### 2022-01-19 src/figure-temperature-logs.R

@@ -114,7 +114,7 @@ cowplot::plot_grid(f1, f2, f3, ncol = 1) -> fig
 ### Save figure
 
 ggsave(fig, file = "results/figures/temperature-logs.png", 
-       path = NULL, scale = 1, width = 182, height = 180, units = "mm", dpi = 600)
+       path = NULL, scale = 1, width = 182, height = 182, units = "mm", dpi = 600)
 # ggsave(fig, file = "results/figures/temperature-logs.tiff", device = grDevices::tiff, 
 #        path = NULL, scale = 1, width = 182, height = 160, units = "mm", dpi = 600, compression = "lzw")
 

### 2022-01-19 src/indices.R

@@ -1,0 +1,71 @@
+library(tidyverse)
+
+### Temporal survey
+
+read.csv("data/temporal-survey-temperatures.csv") %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
+  mutate(Month = lubridate::month(Time)) %>% 
+  mutate(Day = lubridate::day(Time)) %>% 
+  mutate(Hour = lubridate::hour(Time)) %>% 
+  filter(Hour %in% c(0, 4, 8, 12, 16, 20)) %>% # Keep same recording hours as ibuttons
+  filter(! Month %in% 9) %>% # Remove missing September days
+  filter(! (Month %in% 8 & Day > 6)) %>% # Remove missing August days
+  filter(! (Month %in% 10 & Day < 3)) %>% # Remove missing October days %>%
+  dplyr::select(-c(Month, Day, Hour)) %>%
+  group_by(Site, Day = lubridate::floor_date(Time, "day")) %>%
+  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
+  mutate(Snow = ifelse(X < 0.5 & N > -0.5, 1, 0)) %>% # Day is snow day or not
+  # mutate(Snow = ifelse((X - N) <= 0.5, 1, 0)) %>% # Day is snow day or not
+  mutate(FreezeThaw = ifelse(X > 0.5 & N < -0.5, 1, 0)) %>% # Day with freeze-thaw cycles
+  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
+  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
+  group_by(Site, Month = lubridate::floor_date(Day, "month")) %>%
+  summarise(T = mean(T), X = mean(X), N = mean(N), # Daily mean, max, min
+            Snow = sum(Snow), # Snow days per month
+            FreezeThaw = sum(FreezeThaw), # Freeze-thaw days per month
+            FDD = sum(FDD), # FDD per month
+            GDD = sum(GDD)) %>% # GDD per month
+  group_by(Site, Year = lubridate::floor_date(Month, "year")) %>%
+  summarise(bio1 = mean(T), # Annual Mean Temperature
[…]

### 2022-01-20 doc/manuscript.Rmd
new nmds
species chorology

@@
 knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r}
+library(vegan)
+load("../results/numerical/environmental-fit.R")
+load("../results/numerical/bioclim-correlations.R")
 ```
@@
 The main calcareous massif occupies c.
-50 km2 and support high diversity of ecosystems, with alpine vegetation mostly occurring between 1900 and 2400 m a.s.l., with a local species pool of XX species (Jiménez-Alfaro et al.
+50 km^2^ and support high diversity of ecosystems, with alpine vegetation mostly occurring between 1900 and 2400 m a.s.l., with a local species pool of XX species (Jiménez-Alfaro et al.
 XXX).
@@
 We selected four study sites along a north-south gradient, reflecting the main macroclimatic variation from oceanic to Mediterranean influence.
-In 2016, we additionally installed a Water potential logger (XXX) to characterize water stress in the four sites and to ....
 The sites were selected to...
@@
 (Borja)
-
-> Empiezo a pensar que sería mejor quitar esto, por mucho que me fastidie, y guardarnos el tema del WP para otro paper junto con lo de Omaña.
-Eso simplificaría el gráfico (de otro modo, tendría más lógica ponerlo con la Figura 1, porque en realidad no hacemos nada con esto)
 
@@
 
-In the temporal survey (2 plots per site, n =16), we recorded XX species in 2008 and XX in 2018.
-(species loss, gains)...
+In the temporal survey (2 visits x 2 plots per site, n = 16) we recorded 42 species in 2009 and 45 in 2019.
+Of the species recorded in 2009, we did not find again in 2019 the following three: *Festuca burnatii*, *Galium pyrenaicum* and *Viola riviniana*.
+Conversely, in 2019 we recorded six species that we had not seen in 2009: *Arenaria purpurascens*, *Gentiana verna*, *Potentilla crantzii*, *Sedum brevifolium* and *Seseli montanum*.
 Changes in species frequency were/were not correlated with XX and XX....
 (Table 1) (Eduardo).
-In the spatial survey (20 plots per site, n = 80; including four plots placed in rocks with no plants) we recorded XX species.
-Changes in species composition were associated to GDD (NMDS axis 1) to represent a main gradient from the warmest to the coldest sites (Figure 4), and to FDD (NMDS2) in a gradient of snow versus frost winters.
+In the spatial survey (20 plots per site, n = 80; including one plot placed in rocks with no vascular plants) we recorded 81 species, 37 of which were absent from the temporal survey plots.
+Across the whole study system, the five most frequent species were *Thymus praecox* subsp.
+*ligusticus* (83 occurences), *Anthyllis vulneraria* (73), *Koeleria vallesiana* (59), *Minuartia verna* (55) and *Helianthemum canum* (52).
+
+Changes in species composition (**Fig.
+4**) were associated to GDD (NMDS axis 1) to represent a main gradient from the warmest to the coldest microsites, and to FDD (NMDS2) in a gradient of snow versus frost winters.
 In the compositional space, the sites were poorly differentiated across the four sites (PERMANOVA, adonis) mainly by the effect of one site (Los Boches) with most plots places in the cold and frost margins.
@@
 
+Results of the environmental fit:
+
+```{r}
+print(ef1)
+```
+
+Correlations between bioclimatic variables:
+
+```{r}
+print(biocor)
+```
+
+Let us keep GDD, bio7 and FDD for plotting NMDS (low autocorrelation, high r2 in environmental fit)
+
 ## Microclimatic ranges in time and space

### 2022-01-20 src/species-numbers.R

@@ -1,0 +1,39 @@
+library(tidyverse)
+
+read.csv("data/spatial-survey-species.csv") %>% pull(Taxon) %>% unique %>% length 
+read.csv("data/temporal-survey-species.csv") %>% pull(Taxon) %>% unique %>% length 
+
+read.csv("data/spatial-survey-species.csv") %>% select(Taxon) %>%
+  filter(! Taxon %in% (read.csv("data/temporal-survey-species.csv") %>% pull(Taxon) %>% unique)) %>%
+  pull(Taxon) %>% unique %>% length 
+
+read.csv("data/temporal-survey-species.csv") %>%
+  separate(Plot, into = c("Plot", "Year")) %>%
+  select(Taxon, Year) %>%
+  unique %>%
+  group_by(Year) %>% 
+  tally()
+
+
+read.csv("data/temporal-survey-species.csv") %>%
+  separate(Plot, into = c("Plot", "Year")) %>%
+  select(Taxon, Year) %>%
+  unique %>%
+  mutate(Value = 1) %>%
+  spread(Year, Value) %>%
+  filter(is.na(`19`))
+
+read.csv("data/temporal-survey-species.csv") %>%
+  separate(Plot, into = c("Plot", "Year")) %>%
+  select(Taxon, Year) %>%
+  unique %>%
+  mutate(Value = 1) %>%
+  spread(Year, Value) %>%
+  filter(is.na(`09`))
+
+read.csv("data/spatial-survey-species.csv") %>% pull(Plot) %>% unique %>% length
+
+rbind(read.csv("data/spatial-survey-species.csv"), read.csv("data/temporal-survey-species.csv")) %>%
+  select(Taxon, Plot) %>%
+  unique %>%
+  group_by(Taxon) %>% tally %>% arrange(-n)

### 2022-01-20 doc/manuscript.Rmd
species changes

@@
 
-The study was conducted in the calcareous massif of Picos de Europa National Park, in northern Spain (Figure 1A).
+The study was conducted in the central calcareous massif of the Picos de Europa National Park, in northern Spain (Figure 1A).
 The study area is a biodiversity hotspot for cold-adapted plants in the Iberian Peninsula and a biogeographical hub for Alpine and Mediterranean lineages in Western Europe (XXX).
-The main calcareous massif occupies c.
+The central calcareous massif occupies c.
 50 km^2^ and support high diversity of ecosystems, with alpine vegetation mostly occurring between 1900 and 2400 m a.s.l., with a local species pool of XX species (Jiménez-Alfaro et al.
@@
 
-In the temporal survey (2 visits x 2 plots per site, n = 16) we recorded 42 species in 2009 and 45 in 2019.
-Of the species recorded in 2009, we did not find again in 2019 the following three: *Festuca burnatii*, *Galium pyrenaicum* and *Viola riviniana*.
-Conversely, in 2019 we recorded six species that we had not seen in 2009: *Arenaria purpurascens*, *Gentiana verna*, *Potentilla crantzii*, *Sedum brevifolium* and *Seseli montanum*.
+Across the whole study system (temporal and spatial surveys) we recorded 86 species of vascular plants (considering *Helianthemum apenninum* subsp.
+*urrielense* and *Helianthemum apenninum* subsp.
+*cantabricum* as separate species), representing % of the total species pool of the study region.
+The five most frequent species were *Thymus praecox* subsp.
+*ligusticus* (83 occurences), *Anthyllis vulneraria* (73), *Koeleria vallesiana* (59), *Minuartia verna* (55) and *Helianthemum canum* (52).
+Average species richness per 1m^2^ plot was 13, with the richest plot having 25 species and the poorest one just two.
+
+In the temporal survey (2 visits x 2 plots per site, n = 16) we recorded 42 species in 2009 and 47 in 2019.
+Of the species recorded in 2009, we did not find again in 2019 the following three: *Festuca burnatii*, *Galium pyrenaicum* and *Iberis carnosa*.
+Conversely, in 2019 we recorded eight species that we had not seen in 2009: *Arenaria purpurascens*, *Lotus corniculatus*, *Potentilla crantzii*, *Sedum album*, *Sedum brevifolium*, *Seseli montanum*, *Silene ciliata* and *Solidago virgaurea*.
+Of these species that appeared or disappeared from the temporal survey plots, *Arenaria purpurascens* appeared in 12 10x10 cm cells in 2019, while the rest of the cases amounted to less than 10 cells.
+The five species with the highest decrease in frequency from 2009 to 2019 (ignoring annual species and species that occurred in less than 10 10x10 cm cells in 2009) were *Armeria cantabrica* (85% decrease in frequency, present in 13 cells in 2009), *Poa alpina* (-83%, 18 cells), *Salix breviserrata* (-48%, 25 cells), *Jurinea humilis* (-26%, 23 cells) and *Ranunculus parnassiifolius* subsp.
+*favargeri* (-18%, 72 cells).
+The five species with the highest increases (again, ignoring annual species and species that occurred in less than 10 10x10 cm cells in 2009) were *Minuartia verna* (+278%, 19 cells), *Helianthemum apenninum* subsp.
+*urrielense* (+87%, 63 cells), *Arenaria moehringioides* (+85%, 13 cells), *Saxifraga conifera* (+83%, 24 cells) and *Silene acaulis* (+39%, 18 cells).
+
 Changes in species frequency were/were not correlated with XX and XX....
 (Table 1) (Eduardo).
-In the spatial survey (20 plots per site, n = 80; including one plot placed in rocks with no vascular plants) we recorded 81 species, 37 of which were absent from the temporal survey plots.
-Across the whole study system, the five most frequent species were *Thymus praecox* subsp.
-*ligusticus* (83 occurences), *Anthyllis vulneraria* (73), *Koeleria vallesiana* (59), *Minuartia verna* (55) and *Helianthemum canum* (52).
+In the spatial survey (20 plots per site, n = 80; including one plot placed in rocks with no vascular plants) we recorded 81 species, 36 of which were absent from the temporal survey plots.
 

### 2022-01-20 src/indices.R

@@ -69,3 +69,28 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
 rbind(dfSpace, dfTime) -> bioclim
 
 bioclim %>% write.csv("results/numerical/indices.csv", row.names = FALSE)
+
+### SNCs
+
+read.csv("data/spatial-survey-species.csv") %>%
+  merge(bioclim, by = "Plot") %>%
+  select(-Survey) -> o1
+ 
+read.csv("data/temporal-survey-header.csv") %>%
+  select(Plot, Site) %>%
+  merge(read.csv("data/temporal-survey-species.csv")) -> o2
+
+bioclim %>%
+  filter(Survey == "Temporal") %>%
+  gather(Trait, Value, bio1:GDD) %>%
+  group_by(Site, Trait) %>%
+  summarise(Value = mean(Value)) %>%
+  spread(Trait, Value) %>%
+  merge(o2) -> o3
+
+rbind(o1, o3) %>%
+  gather(Trait, Value, bio1:GDD) %>%
+  group_by(Taxon, Trait) %>%
+  summarise(Value = weighted.mean(Value, Cover), n = length(Plot)) %>%
+  spread(Trait, Value) %>%
+  write.csv("results/numerical/snc.csv", row.names = FALSE)

### 2022-01-20 src/model-spp-changes.R

@@ -1,0 +1,50 @@
+library(tidyverse)
+
+### Species traits
+
+read.csv("data/species-checklist.csv") %>%
+  merge(read.csv("results/numerical/snc.csv")) -> traits
+
+traits %>% group_by(Chorology) %>% summarise(GDD = mean(GDD), FDD = mean(FDD), T = mean(T))
+
+### Species changes
+
+read.csv("data/temporal-survey-matrices.csv") %>%
+  group_by(Plot, Taxon) %>%
+  summarise(Cover = sum(Present)) %>%
+  merge(read.csv("data/species-checklist.csv"), all.x = TRUE) %>%
+  filter(Lifeform != "Therophyte") %>%
+  separate(Plot, into = c("Plot", "Year"), sep = "-") %>%
+  mutate(Year = fct_recode(Year, "09" = "08")) %>%
+  group_by(Taxon, Year) %>%
+  summarise(Cover = sum(Cover)) %>%
+  spread(Year, Cover) %>%
+  mutate(Change = `19` - `09`) %>%
+  mutate(ChangeP = Change / `09`) %>%
+  arrange(Change) %>%
+  na.omit %>%
+  merge(read.csv("data/species-checklist.csv")) %>%
+  merge(read.csv("results/numerical/snc.csv")) -> species
+
+### Model total change
+
+lm(Change ~ Chorology + Lifeform + GDD + FDD, data = species) -> m1
+
+summary(m1)
+
+MASS::stepAIC(m1) -> m2
+
+summary(m2)
+
+### Model % change
+
+lm(ChangeP ~ Chorology + Lifeform + GDD + FDD + `09`, data = species) -> m1
+
+summary(m1)
+
+MASS::stepAIC(m1) -> m2
+
+summary(m2)
+
+plot(m2)
+

### 2022-01-20 src/species-numbers.R

@@ -1,13 +1,13 @@
 library(tidyverse)
 
 read.csv("data/spatial-survey-species.csv") %>% pull(Taxon) %>% unique %>% length 
-read.csv("data/temporal-survey-species.csv") %>% pull(Taxon) %>% unique %>% length 
+read.csv("data/temporal-survey-matrices.csv") %>% pull(Taxon) %>% unique %>% length 
 
 read.csv("data/spatial-survey-species.csv") %>% select(Taxon) %>%
-  filter(! Taxon %in% (read.csv("data/temporal-survey-species.csv") %>% pull(Taxon) %>% unique)) %>%
+  filter(! Taxon %in% (read.csv("data/temporal-survey-matrices.csv") %>% pull(Taxon) %>% unique)) %>%
   pull(Taxon) %>% unique %>% length 
 
-read.csv("data/temporal-survey-species.csv") %>%
+read.csv("data/temporal-survey-matrices.csv") %>%
   separate(Plot, into = c("Plot", "Year")) %>%
   select(Taxon, Year) %>%
   unique %>%
@@ -15,20 +15,20 @@ read.csv("data/temporal-survey-species.csv") %>%
   tally()
 
 
-read.csv("data/temporal-survey-species.csv") %>%
+read.csv("data/temporal-survey-matrices.csv") %>%
   separate(Plot, into = c("Plot", "Year")) %>%
-  select(Taxon, Year) %>%
+  group_by(Taxon, Year) %>%
+  tally %>%
   unique %>%
-  mutate(Value = 1) %>%
-  spread(Year, Value) %>%
+  spread(Year, n) %>%
   filter(is.na(`19`))
 
-read.csv("data/temporal-survey-species.csv") %>%
+read.csv("data/temporal-survey-matrices.csv") %>%
   separate(Plot, into = c("Plot", "Year")) %>%
-  select(Taxon, Year) %>%
+  group_by(Taxon, Year) %>%
+  tally %>%
   unique %>%
-  mutate(Value = 1) %>%
-  spread(Year, Value) %>%
+  spread(Year, n) %>%
   filter(is.na(`09`))
 
 read.csv("data/spatial-survey-species.csv") %>% pull(Plot) %>% unique %>% length
@@ -37,3 +37,13 @@ rbind(read.csv("data/spatial-survey-species.csv"), read.csv("data/temporal-surve
   select(Taxon, Plot) %>%
   unique %>%
[…]

### 2022-02-15 doc/manuscript.Rmd
minor changes ms

@@
 title: "Microclimatic spatial variation buffers local warming in alpine vegetation"
-output: github_document
+output: html_document
 bibliography: refs.bib
@@
 .... (Eduardo) (quizás repensar esta parte, pero es importante mantenerla porque es la clave).
+
+## Figures
+
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Study system."}
+knitr::include_graphics("../results/figures/survey-sites.png")
+```
+
+```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Study system."}
+knitr::include_graphics("../results/figures/temperature-logs.png")
+```
+
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Study system."}
+knitr::include_graphics("../results/figures/pca-temperatures.png")
+```
+
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Study system."}
+knitr::include_graphics("../results/figures/nmds-species.png")
+```
+
+```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Study system."}
+knitr::include_graphics("../results/figures/spp-changes.png")
+```

### 2022-02-15 src/species-numbers.R

@@ -1,13 +1,13 @@
 library(tidyverse)
 
 read.csv("data/spatial-survey-species.csv") %>% pull(Taxon) %>% unique %>% length 
-read.csv("data/temporal-survey-matrices.csv") %>% pull(Taxon) %>% unique %>% length 
+read.csv("data/temporal-survey-species.csv") %>% pull(Taxon) %>% unique %>% length 
 
 read.csv("data/spatial-survey-species.csv") %>% select(Taxon) %>%
-  filter(! Taxon %in% (read.csv("data/temporal-survey-matrices.csv") %>% pull(Taxon) %>% unique)) %>%
+  filter(! Taxon %in% (read.csv("data/temporal-survey-species.csv") %>% pull(Taxon) %>% unique)) %>%
   pull(Taxon) %>% unique %>% length 
 
-read.csv("data/temporal-survey-matrices.csv") %>%
+read.csv("data/temporal-survey-species.csv") %>%
   separate(Plot, into = c("Plot", "Year")) %>%
   select(Taxon, Year) %>%
   unique %>%
@@ -15,20 +15,20 @@ read.csv("data/temporal-survey-matrices.csv") %>%
   tally()
 
 
-read.csv("data/temporal-survey-matrices.csv") %>%
+read.csv("data/temporal-survey-species.csv") %>%
   separate(Plot, into = c("Plot", "Year")) %>%
-  group_by(Taxon, Year) %>%
-  tally %>%
+  select(Taxon, Year) %>%
   unique %>%
-  spread(Year, n) %>%
+  mutate(Value = 1) %>%
+  spread(Year, Value) %>%
   filter(is.na(`19`))
 
-read.csv("data/temporal-survey-matrices.csv") %>%
+read.csv("data/temporal-survey-species.csv") %>%
   separate(Plot, into = c("Plot", "Year")) %>%
-  group_by(Taxon, Year) %>%
-  tally %>%
+  select(Taxon, Year) %>%
   unique %>%
-  spread(Year, n) %>%
+  mutate(Value = 1) %>%
+  spread(Year, Value) %>%
   filter(is.na(`09`))
 
 read.csv("data/spatial-survey-species.csv") %>% pull(Plot) %>% unique %>% length
@@ -37,13 +37,3 @@ rbind(read.csv("data/spatial-survey-species.csv"), read.csv("data/temporal-surve
   select(Taxon, Plot) %>%
   unique %>%
[…]

### 2022-02-16 doc/manuscript.Rmd
glm species models 🔮

@@
 ```{r}
-library(vegan)
+library(vegan); library(tidyverse)
 load("../results/numerical/environmental-fit.R")
@@
 ```
+
+## Tables
+
+```{r echo = FALSE, fig.cap = "**Figure 1** Study system."}
+read.csv("../results/tables/T1 - Summary of the GLM models of species presence.csv", check.names = FALSE) %>%
+  kableExtra::kbl(caption = "Table 1 Summary of the GLM models of species presence") %>%
+  kableExtra::kable_classic(., full_width = F)
+```

### 2022-02-16 src/glms.R

@@ -1,0 +1,139 @@
+library(tidyverse)
+
+### Prepare data
+
+read.csv("data/spatial-survey-species.csv") %>%
+  group_by(Taxon) %>%
+  mutate(n = length(Taxon)) %>%
+  filter(n >= 10) %>%
+  mutate(Presence = 1) %>%
+  select(Plot, Taxon, Presence) %>%
+  spread(Taxon, Presence, fill = 0) %>%
+  gather(Taxon, Presence, -Plot) %>%
+  merge(read.csv("results/numerical/indices.csv")) -> presences
+
+### Get GLM coefficients
+
+glms <- function(x) {
+  glm(Presence ~ GDD + FDD, family = "binomial", data = x) -> m1
+  broom::tidy(m1)
+}
+
+presences %>%
+  group_by(Taxon) %>%
+  do(glms(.)) %>%
+  gather(Trait, Value, estimate:p.value) %>%
+  mutate(Trait = paste(term, Trait)) %>%
+  select(Taxon, Trait, Value) %>%
+  spread(Trait, Value) %>%
+  select(`Taxon`, 
+         `(Intercept) estimate`, `(Intercept) std.error`, `(Intercept) statistic`, `(Intercept) p.value`, 
+         `FDD estimate`, `FDD std.error`, `FDD statistic`, `FDD p.value`, 
+         `GDD estimate`, `GDD std.error`, `GDD statistic`, `GDD p.value`) -> t1
+
+### Get GLM summary statistics
+
+glms.glance <- function(x) {
+  glm(Presence ~ GDD + FDD, family = "binomial", data = x) -> m1
+  broom::glance(m1)
+}
+
+presences %>%
+  group_by(Taxon) %>%
+  do(glms.glance(.)) -> t2
+
+### Get McFadden's pseudo R2 (rho2) https://stats.stackexchange.com/questions/82105/mcfaddens-pseudo-r2-interpretation
+
+glms.mcfadden <- function(x) { 
+  glm(Presence ~ GDD + FDD, family = "binomial", data = x) -> mod
+  glm(Presence ~ 1, family = "binomial", data = x) -> nullmod
+  data.frame(rho2 = 1 - logLik(mod) / logLik(nullmod)) #https://thestatsgeek.com/2014/02/08/r-squared-in-logistic-regression/
+}
+
+presences %>%
+  group_by(Taxon) %>%
+  do(glms.mcfadden(.)) %>%
[…]

### 2022-02-16 src/figure-nmds-species.R
species x-tinctions 💀

@@ -103,10 +103,10 @@ ggplot(sitescores, aes(x = Dim.1, y = Dim.2)) +
   scale_x_continuous(name = "NMDS1") + 
   scale_y_continuous(name = "NMDS2") +
   scale_fill_manual(values = c("goldenrod", "forestgreen",  "royalblue", "darkorchid")) +
-  annotate("label", x = -.8, y = -1, label = "Warm & Snow", fill = "indianred", color = "white", size = 4) +
-  annotate("label", x = .7, y = -1, label = "Cold & Snow", fill = "deepskyblue", color = "white", size = 4) +
-  annotate("label", x = -.8, y = 1, label = "Warm & Frost", fill = "firebrick", color = "white", size = 4) +
-  annotate("label", x = .7, y = 1, label = "Cold & Frost", fill = "slateblue3", color = "white", size = 4) -> f2; f2
+  annotate("label", x = -.8, y = -1, label = "Hot & Snowy", fill = "indianred", color = "white", size = 4) +
+  annotate("label", x = .7, y = -1, label = "Cold & Snowy", fill = "deepskyblue", color = "white", size = 4) +
+  annotate("label", x = -.8, y = 1, label = "Hot & Frozen", fill = "firebrick", color = "white", size = 4) +
+  annotate("label", x = .7, y = 1, label = "Cold & Frozen", fill = "slateblue3", color = "white", size = 4) -> f2; f2
 
 ### Save figure
 

### 2022-02-16 src/glms.R

@@ -121,7 +121,7 @@ t1 %>%
 
 write.csv(models, "results/supplement/S3 - Output of the GLM models of species presence.csv", row.names = FALSE)
 
-### Figure
+### Table 1 - GLMS
 
 models %>%
   filter(`FDD p.value` < 0.05 | `GDD p.value` < 0.05) %>%
@@ -137,3 +137,16 @@ models %>%
   select(Taxon, `GDD estimate`, `GDD p`, `FDD estimate`, `FDD p`,
          rho2, `Cold & Frozen`:`Hot & Snowy`, `Frequency 2009`, `Frequency 2018`) %>%
   write.csv("results/tables/T1 - Summary of the GLM models of species presence.csv", row.names = FALSE)
+
+### Figure 5 extinctions
+
+models %>%
+  filter(`FDD p.value` < 0.05 | `GDD p.value` < 0.05) %>%
+  filter(rho2 > 0.15) %>%
+  select(Taxon, `Cold & Frozen`:`Hot & Snowy`) %>%
+  gather(Scenario, Predict, `Cold & Frozen`:`Hot & Snowy`) %>%
+  filter(Predict == 0) %>%
+  group_by(Scenario) %>%
+  tally %>%
+  ggplot(aes(reorder(Scenario, n), n)) +
+  geom_bar(stat = "identity", position = "dodge")

### 2022-02-16 src/indices.R

@@ -68,29 +68,4 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
 
 rbind(dfSpace, dfTime) -> bioclim
 
-bioclim %>% write.csv("results/numerical/indices.csv", row.names = FALSE)
-
-### SNCs
-
-read.csv("data/spatial-survey-species.csv") %>%
-  merge(bioclim, by = "Plot") %>%
-  select(-Survey) -> o1
- 
-read.csv("data/temporal-survey-header.csv") %>%
-  select(Plot, Site) %>%
-  merge(read.csv("data/temporal-survey-species.csv")) -> o2
-
-bioclim %>%
-  filter(Survey == "Temporal") %>%
-  gather(Trait, Value, bio1:GDD) %>%
-  group_by(Site, Trait) %>%
-  summarise(Value = mean(Value)) %>%
-  spread(Trait, Value) %>%
-  merge(o2) -> o3
-
-rbind(o1, o3) %>%
-  gather(Trait, Value, bio1:GDD) %>%
-  group_by(Taxon, Trait) %>%
-  summarise(Value = weighted.mean(Value, Cover), n = length(Plot)) %>%
-  spread(Trait, Value) %>%
-  write.csv("results/numerical/snc.csv", row.names = FALSE)
+bioclim %>% write.csv("results/supplement/S1 - Bioclimatic indices.csv", row.names = FALSE)

### 2022-02-16 src/glms.R
species x-tinctions 💀

@@ -10,7 +10,7 @@ read.csv("data/spatial-survey-species.csv") %>%
   select(Plot, Taxon, Presence) %>%
   spread(Taxon, Presence, fill = 0) %>%
   gather(Taxon, Presence, -Plot) %>%
-  merge(read.csv("results/numerical/indices.csv")) -> presences
+  merge(read.csv("results/supplement/S1 - Bioclimatic indices.csv")) -> presences
 
 ### Get GLM coefficients
 
@@ -58,12 +58,12 @@ presences %>%
 ### Scenarios
 
 merge(
-read.csv("results/numerical/indices.csv") %>%
+read.csv("results/supplement/S1 - Bioclimatic indices.csv") %>%
   select(Survey, GDD) %>%
   filter(Survey == "Temporal") %>%
   # group_by(Survey) %>%
   filter(GDD == max(GDD) | GDD == min(GDD)),
-read.csv("results/numerical/indices.csv") %>%
+read.csv("results/supplement/S1 - Bioclimatic indices.csv") %>%
   select(Survey, FDD) %>%
   filter(Survey == "Temporal") %>%
   # group_by(Survey) %>%
@@ -148,5 +148,64 @@ models %>%
   filter(Predict == 0) %>%
   group_by(Scenario) %>%
   tally %>%
-  ggplot(aes(reorder(Scenario, n), n)) +
-  geom_bar(stat = "identity", position = "dodge")
+  mutate(Scenario = gsub(" & ", "\n", Scenario)) %>%
+  ggplot(aes(reorder(Scenario, n), n, fill = Scenario)) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  geom_bar(stat = "identity", position = "dodge", color = "black") +
+  labs(title = "(A) Species extinctions per scenario") +
+  xlab("Scenario") +
+  ylab("Number of species extinctions") +
+  ggthemes::theme_tufte() +
+  scale_fill_manual(values = c("slateblue3", "deepskyblue", "firebrick", "indianred")) +
+  scale_y_continuous(breaks = seq(0, 10, by = 1)) +
+  theme(legend.position = "none", 
+        legend.direction = "vertical",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 16, face = "italic"), 
+        panel.background = element_rect(color = "grey96", fill = "grey96"),
+        strip.text = element_text(size = 12),
+        strip.placement = "ouside",
+        axis.title.x = element_text(size = 12, color = "black"),
[…]

### 2022-02-21 src/figure-pca-temperatures.R
1st full draft of figures 🖼️

@@ -1,99 +1,0 @@
-library(tidyverse)
-
-read.csv("results/numerical/indices.csv") -> bioclim
-bioclim %>% filter(Site == "Los Cazadores") -> bioclim1
-bioclim %>% filter(Site == "Hou Sin Tierri") -> bioclim2
-bioclim %>% filter(Site == "Los Boches") -> bioclim3
-bioclim %>% filter(Site == "Hoyo Sin Tierra") -> bioclim4
-
-### PCA 1
-
-bioclim1[, 3:8] %>%
-  FactoMineR::PCA() -> pca1
-
-cbind((bioclim1 %>%  dplyr::select(Plot, Site, Survey)), data.frame(pca1$ind$coord[, 1:2])) -> pcaInds1
-
-pca1$var$coord[, 1:2] %>%
-  data.frame %>%
-  rownames_to_column(var = "Variable") %>%
-  mutate(Site = "Los Cazadores") -> pcaVars1
-
-### PCA 2
-
-bioclim2[, 3:8] %>%
-  FactoMineR::PCA() -> pca2
-
-cbind((bioclim2 %>%  dplyr::select(Plot, Site, Survey)), data.frame(pca2$ind$coord[, 1:2])) -> pcaInds2
-
-pca2$var$coord[, 1:2] %>%
-  data.frame %>%
-  rownames_to_column(var = "Variable") %>%
-  mutate(Site = "Hou Sin Tierri") -> pcaVars2
-
-### PCA 3
-
-bioclim3[, 3:8] %>%
-  FactoMineR::PCA() -> pca3
-
-cbind((bioclim3 %>%  dplyr::select(Plot, Site, Survey)), data.frame(pca3$ind$coord[, 1:2])) -> pcaInds3
-
-pca3$var$coord[, 1:2] %>%
-  data.frame %>%
-  rownames_to_column(var = "Variable") %>%
-  mutate(Site = "Los Boches") -> pcaVars3
-
-### PCA 4
-
-bioclim4[, 3:8] %>%
-  FactoMineR::PCA() -> pca4
-
-cbind((bioclim4 %>%  dplyr::select(Plot, Site, Survey)), data.frame(pca4$ind$coord[, 1:2])) -> pcaInds4
-
-pca4$var$coord[, 1:2] %>%
-  data.frame %>%
-  rownames_to_column(var = "Variable") %>%
-  mutate(Site = "Hoyo Sin Tierra") -> pcaVars4
-
[…]

### 2022-02-21 src/figure-spp-changes.R

@@ -1,90 +1,0 @@
-library(tidyverse)
-
-### NMDS species scores
-
-read.csv("data/spatial-survey-species.csv") %>%
-  filter(Plot != "A03") %>%
-  filter(Plot != "D19") %>%
-  spread(Taxon, Cover, fill = 0) %>%
-  arrange(Plot) %>%
-  column_to_rownames(var = "Plot") %>%
-  vegan::metaMDS(trymax = 100000, k = 2) %>%
-  vegan::scores(., "species") %>%
-  data.frame() %>%
-  rownames_to_column("Taxon") %>%
-  rename(Dim.1 = NMDS1, Dim.2 = NMDS2) %>%
-  mutate(Species = gsub(" gr\\.", "", Taxon)) %>%
-  mutate(Species = gsub("apenninum subsp. ", "", Species)) %>%
-  separate(Species, into = c("G", "S")) %>%
-  mutate(Species = paste(substr(G, 1, 3), substr(S, 1, 3), sep = "")) %>%
-  dplyr::select(Taxon, Species, Dim.1, Dim.2) -> sppscores
-
-### SNCs
-
-read.csv("data/spatial-survey-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
-  group_by(Plot, Day = lubridate::floor_date(Time, "day")) %>%
-  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
-  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
-  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
-  group_by(Plot, Month = lubridate::floor_date(Day, "month")) %>%
-  summarise(FDD = sum(FDD), # FDD per month
-            GDD = sum(GDD)) %>% # GDD per month
-  group_by(Plot) %>%
-  summarise(FDD = sum(FDD), # FDD per year
-            GDD = sum(GDD))  %>%
-  select(Plot, FDD, GDD) %>%
-  merge(read.csv("data/spatial-survey-species.csv")) %>%
-  group_by(Taxon) %>%
-  summarise(n = length(Plot),
-            GDD = weighted.mean(GDD, Cover),
[…]

### 2022-02-21 src/figure-temperature-deviations.R

@@ -1,90 +1,0 @@
-library(tidyverse); library(lubridate)
-
-### Temporal survey data
-
-read.csv("data/temporal-survey-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
-  group_by(Site, Day = lubridate::floor_date(Time, "day")) %>%
-  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>%
-  mutate(d = day(Day), 
-         m = month(Day)) -> years
-
-years %>%
-  group_by(Site, m, d) %>%
-  summarise(TM = mean(T)) %>%
-  merge(years) %>% 
-  mutate(Site = fct_relevel(Site,
-                            "Los Cazadores", "Hou Sin Tierri",
-                            "Los Boches", "Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> time
-
-### Spatial survey data
-
-read.csv("data/spatial-survey-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
-  merge(read.csv("data/spatial-survey-header.csv")) %>%
-  group_by(Plot, Site, Day = lubridate::floor_date(Time, "day")) %>%
-  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) -> sites
-
-sites %>%
-  group_by(Site, Day) %>%
-  summarise(TM = mean(T)) %>%
-  merge(sites) %>% 
-  mutate(Site = fct_relevel(Site,
-                            "Los Cazadores", "Hou Sin Tierri",
-                            "Los Boches", "Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> space
-
-### Temporal survey deviations
-
-time %>%
-  ggplot(aes(TM, T, fill = Site)) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  geom_vline(xintercept = 0, linetype = "dashed") +
-  geom_point(alpha = 0.5, shape = 21) +
-  geom_abline(intercept = 0, slope = 1, size = 1) +
-  scale_fill_manual(values = c("goldenrod", "forestgreen",  "royalblue", "darkorchid")) +
-  ggthemes::theme_tufte() + 
[…]

### 2022-02-21 src/figures-spacevstime.R

@@ -1,168 +1,0 @@
-library(tidyverse); library(tibbletime); library(vegan)
-
-### Indices for GeoPrecision by complete year
-
-read.csv("data/temporal-survey-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
-  group_by(Site, Day = lubridate::floor_date(Time, "day")) %>%
-  summarise(T = mean(Temperature), X = max(Temperature), N = min(Temperature), n = length(Time)) %>% # Daily mean, max, min
-  mutate(Snow = ifelse(X < 0.5 & N > -0.5, 1, 0)) %>% # Day is snow day or not
-  # mutate(Snow = ifelse((X - N) <= 0.5, 1, 0)) %>% # Day is snow day or not
-  mutate(FreezeThaw = ifelse(X > 0.5 & N < -0.5, 1, 0)) %>% # Day with freeze-thaw cycles
-  mutate(FDD = ifelse(T < 0, T, 0)) %>% # Freezing degrees per day
-  mutate(GDD = ifelse(T >= 1, T, 0)) %>% # Growing degrees day per month https://link.springer.com/article/10.1007/s00035-021-00250-1
-  group_by(Site, Month = lubridate::floor_date(Day, "month")) %>%
-  summarise(T = mean(T), X = mean(X), N = mean(N), # Daily mean, max, min
-            Snow = sum(Snow), # Snow days per month
-            FreezeThaw = sum(FreezeThaw), # Freeze-thaw days per month
-            FDD = sum(FDD), # FDD per month
-            GDD = sum(GDD)) %>% # GDD per month
-  group_by(Site, Year = lubridate::floor_date(Month, "year")) %>%
-  summarise(bio01 = mean(T), # Annual Mean Temperature
-            bio02 = mean(X - N), # Mean Diurnal Range (Mean of monthly (max temp - min temp))
-            bio04 = sd(T) * 100, # Temperature Seasonality (standard deviation ×100)
-            bio05 = max(X), # Max Temperature of Warmest Month
-            bio06 = min(N), # Min Temperature of Coldest Month
-            bio07 = bio05 - bio06, # Temperature Annual Range (BIO5-BIO6)
-            bio03 = (bio02 / bio07) * 100, # Isothermality (BIO2/BIO7) (×100)
-            Snow = sum(Snow), # Days of snow per year
[…]

### 2022-02-21 src/glms.R

@@ -152,7 +152,6 @@ models %>%
   ggplot(aes(reorder(Scenario, n), n, fill = Scenario)) +
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_bar(stat = "identity", position = "dodge", color = "black") +
-  labs(title = "(A) Species extinctions per scenario") +
   xlab("Scenario") +
   ylab("Number of species extinctions") +
   ggthemes::theme_tufte() +
@@ -169,43 +168,10 @@ models %>%
         axis.title.y = element_text(size = 12, color = "black"),
         axis.text.x = element_text(size = 10, color = c("deepskyblue", "slateblue3", "indianred", "firebrick"), 
                                    face = "bold"),
-        axis.text.y = element_text(size = 10, color = "black")) -> f5a; f5a
-
-### Indices trends figure
-
-read.csv("results/supplement/S1 - Bioclimatic indices.csv") %>%
-  filter(Survey == "Temporal") %>%
-  group_by(Plot) %>%
-  summarise(`Freezing degrees-day` = mean(FDD), `Growing degrees-day` = mean(GDD)) %>%
-  gather(Trait, Value, `Growing degrees-day`:`Freezing degrees-day`) %>%
-  mutate(Trait = fct_relevel(Trait, "Growing degrees-day", "Freezing degrees-day")) %>%
-  ggplot(aes(as.numeric(Plot), Value, fill = Trait)) +
-  facet_wrap(~ Trait, nrow = 2, scales = "free_y") +
-  geom_area(alpha = 0.6, color = "black") +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  labs(title = "(B) Trends in GDD and FDD") +
-  xlab("Year") +
-  ylab(expression(paste("Degrees-day (absolute ºC)"))) +
-  ggthemes::theme_tufte() +
-  scale_fill_manual(values = c("gold", "darkorchid")) +
-  scale_x_continuous(breaks = seq(2009, 2018, by = 1)) +
-  theme(strip.background = element_rect(colour = "grey96", fill = "grey96"),
-        legend.position = "none", 
-        legend.direction = "vertical",
-        legend.title = element_blank(),
-        legend.text = element_text(size = 16, face = "italic"), 
-        panel.background = element_rect(color = "grey96", fill = "grey96"),
-        strip.text = element_text(size = 12),
-        strip.placement = "ouside",
[…]

### 2022-02-21 src/logs.R

@@ -14,7 +14,7 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
   facet_wrap(~ Site, nrow = 1) + 
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_line() +
-  labs(title = "(A) Hourly logs, spatial survey (20 iButtons per site, Oct 2018 - Aug 2019)") +
+  labs(title = "(B) Hourly logs, spatial survey (20 iButtons per site, Oct 2018 - Aug 2019)") +
   xlab("Time (4-h recording inverval)") +
   ylab("Temperature (ºC)") +
   ggthemes::theme_tufte() +
@@ -41,7 +41,7 @@ read.csv("data/temporal-survey-temperatures.csv") %>%
   ggplot(aes(Time, Temperature)) + 
   facet_wrap(~ Site, nrow = 1) + 
   geom_line(color = "red") +
-  labs(title = "(B) Hourly logs, temporal survey (2009-2018)") +
+  labs(title = "(A) Hourly logs, temporal survey (2009-2018)") +
   xlab("Time (1-h recording inverval)") +
   ylab("Temperature (ºC)") +
   geom_hline(yintercept = 0, linetype = "dashed") +
@@ -58,63 +58,14 @@ read.csv("data/temporal-survey-temperatures.csv") %>%
         axis.text.x = element_text(size = 8, color = "black"),
         axis.text.y = element_text(size = 11, color = "black")) -> f2; f2
 
-### Decompose time series
-
-decomposer <- function(x) 
-{
-  x %>% 
-    pull(Temperature) %>%
-    ts(start = c(2009, 1), frequency = 24 * 365) -> timeseries
-  
-  stl(timeseries, s.window = "period") -> fit
-  
-  cbind(x$Time, as.data.frame(timeseries), as.data.frame(fit$time.series)) -> df
-  
-  colnames(df) <- c("Time", "Temperature", "Seasonal", "Trend", "Remainder")
-  
-  df
-}
-
-read.csv("data/temporal-survey-temperatures.csv") %>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
-  group_by(Site) %>%
-  do(decomposer(.)) %>%
-  mutate(Site = as.factor(Site)) %>%
-  mutate(Site = fct_relevel(Site,
-                            "Los Cazadores", "Hou Sin Tierri",
-                            "Los Boches", "Hoyo Sin Tierra")) %>%
[…]

### 2022-02-21 src/map.R

@@ -97,7 +97,7 @@ cowplot::plot_grid(plot1, f3, ncol = 2, rel_widths = c(60, 35)) -> fig
 
 ### Save figure
 
-ggsave(fig, file = "results/figures/survey-sites.png", 
+ggsave(fig, file = "results/figures/F1 - Map.png", 
        path = NULL, scale = 1, width = 95, height = 133, units = "mm", dpi = 600)
 # ggsave(fig, file = "results/figures/survey-sites.tiff", device = grDevices::tiff, 
 #        path = NULL, scale = 1, width = 182, height = 133, units = "mm", dpi = 600, compression = "lzw")

### 2022-02-21 src/model-spp-changes.R

@@ -1,50 +1,0 @@
-library(tidyverse)
-
-### Species traits
-
-read.csv("data/species-checklist.csv") %>%
-  merge(read.csv("results/numerical/snc.csv")) -> traits
-
-traits %>% group_by(Chorology) %>% summarise(GDD = mean(GDD), FDD = mean(FDD), T = mean(T))
-
-### Species changes
-
-read.csv("data/temporal-survey-matrices.csv") %>%
-  group_by(Plot, Taxon) %>%
-  summarise(Cover = sum(Present)) %>%
-  merge(read.csv("data/species-checklist.csv"), all.x = TRUE) %>%
-  filter(Lifeform != "Therophyte") %>%
-  separate(Plot, into = c("Plot", "Year"), sep = "-") %>%
-  mutate(Year = fct_recode(Year, "09" = "08")) %>%
-  group_by(Taxon, Year) %>%
-  summarise(Cover = sum(Cover)) %>%
-  spread(Year, Cover) %>%
-  mutate(Change = `19` - `09`) %>%
-  mutate(ChangeP = Change / `09`) %>%
-  arrange(Change) %>%
-  na.omit %>%
-  merge(read.csv("data/species-checklist.csv")) %>%
-  merge(read.csv("results/numerical/snc.csv")) -> species
-
-### Model total change
-
-lm(Change ~ Chorology + Lifeform + GDD + FDD, data = species) -> m1
-
-summary(m1)
-
-MASS::stepAIC(m1) -> m2
-
-summary(m2)
-
-### Model % change
-
-lm(ChangeP ~ Chorology + Lifeform + GDD + FDD + `09`, data = species) -> m1
-
-summary(m1)
-
-MASS::stepAIC(m1) -> m2
-
-summary(m2)
-
-plot(m2)
-

### 2022-02-21 src/nmds.R

@@ -4,27 +4,15 @@ library(tidyverse)
 
 read.csv("data/spatial-survey-species.csv") %>%
   filter(Plot != "A03") %>%
-  filter(Plot != "D19") -> speciesSS # Vegetation data from the spatial survey
-
-read.csv("data/temporal-survey-species.csv")  -> speciesTS # Vegetation data from the temporal survey
- 
-rbind(speciesSS, speciesTS)  %>%
+  filter(Plot != "D19")  %>%
   spread(Taxon, Cover, fill = 0) %>%
   arrange(Plot) %>%
-  column_to_rownames(var = "Plot") -> species # Merge species
-
+  column_to_rownames(var = "Plot") -> species 
 ### Header data
 
-read.csv("data/spatial-survey-header.csv") %>% mutate(Survey = "Spatial") %>%
-  merge(read.csv("results/numerical/indices.csv")) -> headerSS # Header data from the spatial survey
-
-read.csv("results/numerical/indices.csv") %>%
-  filter(Survey == "Temporal") %>%
-  filter(Plot == "2018") %>% 
-  select(-Plot) %>% 
-  merge(read.csv("data/temporal-survey-header.csv")) -> headerTS # Header data from the temporal survey
-
-rbind(headerSS, headerTS) -> header # Merge header
+read.csv("data/spatial-survey-header.csv") %>% 
+  mutate(Survey = "Spatial") %>%
+  merge(read.csv("results/supplement/S1 - Bioclimatic indices.csv")) -> header 
 
 ### Do NMDS
 
@@ -77,7 +65,7 @@ ef1$vectors$arrows %>%
   data.frame() %>%
   dplyr::select(NMDS1, NMDS2) %>%
   rownames_to_column(var = "Variable") %>%
-  filter(Variable %in% c("GDD", "FDD", "bio7")) %>%
+  filter(Variable %in% c("GDD", "FDD")) %>%
   rename(Dim.1 = NMDS1, Dim.2 = NMDS2) -> envvars
 
 ### Plot NMDS sites
@@ -86,13 +74,13 @@ ggplot(sitescores, aes(x = Dim.1, y = Dim.2)) +
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_vline(xintercept = 0, linetype = "dashed") +
   geom_segment(data = envvars, aes(x = 0, y = 0, xend = Dim.1, yend = Dim.2)) +
-  geom_point(aes(fill = Site, shape = Survey, alpha = Survey), color = "black", size = 4) +
[…]

### 2022-02-21 src/space-time.R

@@ -1,0 +1,78 @@
+library(tidyverse)
+
+read.csv("results/supplement/S1 - Bioclimatic indices.csv") -> bioclim
+
+bioclim %>%
+  gather(Trait, Value, bio1:GDD) %>%
+  group_by(Trait, Site, Survey) %>%
+  filter(Trait %in% c("GDD", "FDD")) %>%
+  summarise(Delta = max(Value) - min(Value)) %>%
+  group_by(Trait, Survey) %>%
+  summarise(Mean = mean(Delta), SE = sd(Delta)/sqrt(4)) %>%
+  group_by() %>%
+  mutate(Trait = fct_relevel(Trait, "GDD", "FDD")) %>%
+  ggplot(aes(Survey, Mean, fill = paste(Trait, Survey))) +
+  facet_wrap(~ Trait, scales = "free_y", nrow = 1) +
+  geom_bar(stat = "identity", position = "dodge", color = "black", alpha = 0.9) +
+  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE), width=.2,
+                position=position_dodge(.9)) +
+  xlab("Survey") +
+  ylab(expression(paste("Degrees-day (absolute ºC)"))) +
+  labs(title = "(A) Max differences spatial vs. temporal") +
+  ggthemes::theme_tufte() +
+  scale_fill_manual(values = c("deepskyblue", "deepskyblue", "brown1", "brown1")) +
+  theme(strip.background = element_rect(colour = "grey96", fill = "grey96"),
+        legend.position = "none", 
+        legend.direction = "vertical",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 16, face = "italic"), 
+        panel.background = element_rect(color = "grey96", fill = "grey96"),
+        strip.text = element_text(size = 12),
+        strip.placement = "ouside",
+        axis.title.x = element_text(size = 12, color = "black"),
+        axis.title.y = element_text(size = 12, color = "black"),
+        axis.text.x = element_text(size = 8, color = "black"),
+        axis.text.y = element_text(size = 10, color = "black")) -> f3a; f3a
+
+### Indices trends figure
+
+read.csv("results/supplement/S1 - Bioclimatic indices.csv") %>%
+  filter(Survey == "Temporal") %>%
+  group_by(Plot) %>%
+  summarise(FDD = mean(FDD), GDD = mean(GDD)) %>%
+  gather(Trait, Value, GDD:FDD) %>%
[…]

### 2022-02-23 doc/manuscript.Rmd
methods 1st draft 🔧

--- 2022-01-20 doc/manuscript.Rmd
@@
 
-```{r}
-library(vegan)
+```{r message = FALSE, echo = FALSE, warning = FALSE}
+library(vegan); library(tidyverse)
 load("../results/numerical/environmental-fit.R")
@@
 
+We conducted all analysis with R [@RN2315], and the code for analysis and creation of the figures and manuscript is available at GitHub (see Data Availability Statement).
+
 ## Study system
 
-The study was conducted in the central calcareous massif of the Picos de Europa National Park, in northern Spain (Figure 1A).
+The study was conducted in the central calcareous massif of the Picos de Europa National Park, in northern Spain (**Figure 1A**).
 The study area is a biodiversity hotspot for cold-adapted plants in the Iberian Peninsula and a biogeographical hub for Alpine and Mediterranean lineages in Western Europe (XXX).
 The central calcareous massif occupies c.
-50 km^2^ and support high diversity of ecosystems, with alpine vegetation mostly occurring between 1900 and 2400 m a.s.l., with a local species pool of XX species (Jiménez-Alfaro et al.
+50 km^2^ and supports a high diversity of ecosystems, with alpine vegetation mostly occurring between 1900 and 2400 m a.s.l., with a local species pool of XX species (Jiménez-Alfaro et al.
 XXX).
 In 2008, we established a long-term ecological research program for monitoring soil climate and vegetation change.
-We selected four study sites along a north-south gradient, reflecting the main macroclimatic variation from oceanic to Mediterranean influence.
+We selected four study sites along a north-south gradient (**Figure 1B**), reflecting the main macroclimatic variation from Atlantic to Mediterranean influence.
 The sites were selected to...
@@
 
-In each site, we buried a temperature logger (XXXX) at 5 cm depth in a relatively flat and homogeneous vegetation patch.
-We surveyed the plant community in two replicated plots of 1 m2 separated 1 m from the logger, identifying species composition and estimating relative cover in %.
+*Temporal survey.* In each site, we buried a temperature logger (M-Log5W, GeoPrecision, Ettlingen, Germany; accuracy: +/- 0.1 ºC at 0 ºC, resolution: 0.01 ºC, records each hour) at 5 cm depth in a relatively flat and homogeneous vegetation patch.
+We surveyed the plant community in two replicated plots of 1 m^2^ separated 1 m from the logger, identifying species composition of vascular plants and estimating relative cover in %.
 In each plot, we installed a grid template of 100 microplots (10 x 10 cm each) to sample species frequency according to the standard methodology of GLORIA (XXX).
@@
 The vegetation data from these surveys, together with the soil temperature collected in the four study sites during 10 years, represent the "temporal survey".
-In 2019, we visited the same areas to study the spatial variation of vegetation and microclimate around the previously sampled areas.
-Using the long-term temperature logger as the central point, we additionally placed 20 ibuttons (XXX) in 20 plots of 1 m2 separated 10 m from each other in the four cardinal directions (Figure 1C)..
-etc...
-(Borja)
 
-## Data analysis
+> En este punto podríamos comentar que el long-term monitoring incluye otros cuatro sitios, algo más diferentes, que por ser más diferentes no los consideramos en este estudio, pero que añadimos los datos en supplementario.
+Así, quedan publicados.
 
-[@RN2315]
+*Spatial survey.* In 2018, we visited the same areas to study the spatial variation of vegetation and microclimate around the previously sampled areas.
+Using the long-term temperature logger as the central point, we additionally placed 20 iButtons (Thermochron, iButton, Newbury, UK; accuracy: +/- 0.5 ºC from -10 ºC to +65 ºC, resolution: 0.5 ºC, records each 4 hours) in 20 plots of 1 m^2^ separated 10 m from each other in the four cardinal directions (**Figure 1C**).
+At each of the 20 plots, we identified vascular plants and estimated their relative cover in %.
+In 2019 we came back to download the data of the iButtons.
+These data, together with the vegetation data of the iButton plots, is the "spatial survey".
 
-Vegetation data + Temperature data (Eduardo)
+## Bioclimatic indices
+
+We used the microclimatic data of the temporal and spatial surveys to calculate several bioclimatic indices.
+We did this for each year of the temporal survey (4 sites * 10 years) and each of the 80 spatial plots (4 sites x 20 plots).
+The spatial data series did not cover a full year, and were missing part of August and September.
+Thus, for comparison, we also removed this period from the temporal data before calculating the indices.
+Additionally, for the temporal survey (hourly records), we kept only the same recording hours as in the spatial survey (records each 4 hours).
+
+The bioclimatic indices were the following: (1) bio1 = annual mean temperature; (2) bio2 = mean diurnal range,i.e.
+the mean of the monthly differences between maximum and minimum temperatures; (3) bio7 = temperature annual range; i.e.
+the difference between the maximum temperature of the warmest month and the minimum temperature of the coldest month; (4) snow = the number of days of snow cover, considered to be those days in which the maximum temperature was below 0.5 ºC and the minimum temperature was above -0.5 ºC; (5) GDD = growing degrees-day, i.e.
+the sum of daily mean temperatures for days in which the mean temperature was above 1 ºC; (6) FDD = freezing degrees-day, i.e.
+the sum of daily mean temperatures for days in which the mean temperature was below 0 ºC.
+The indices bio1, bio2 and bio7 follow the definitions of WorldClim [@RN5064].
+For GDD, we used the 1 ºC threshold for the growing season, following Bürli et al.
+[-@RN5065].
+For FDD, we transformed the values from negative to positive, so the interpretation would be easier (i.e.
+higher values equal more freezing).
+The bioclimatic indices are provided as supplementary material **S1**.
+
+Before proceeding with further analysis, we conducted a principal component analysis (PCA) [@RN3166] of the bioclimatic indices for the spatial survey, to identify the main patterns in microclimatic variability.
+Results of the PCA are shown in supplementary material **S2**.
+In summary, the first axis of variation represented a gradient of increasing thermicity, with higher values of GDD, bio1, bio2 and bio7.
+The second axis represented a gradient of increasing freezing, with higher values of FDD.
+Snow length was mostly associated to the third axis.
+For ease of interpretation, we kept only GDD and FDD for further analysis.
+GDD and FDD were not correlated (r = -0.03) while GDD was correlated to bio1 (r = 0.97), bio2 (r = 0.76) and bio7 (r = 0.45).
+
+## Simulation of species extinctions
+
+We modeled species occurrence at the micro-site level as a function of the bioclimatic indices GDD and FDD, using the data of the spatial survey.
+First, we used non-metric multidimensional scaling (NMDS) with environmental fitting [@RN3388] to visualize the patterns of change in vegetation, and their relation to GDD and FDD.
+
+Then, we used Generalized Linear Models (GLMs, binomial family) to model species presence (1 or 0) in each of the spatial plots (n = 78, we removed two plots with no vascular plants) as a response to the plot's GDD and FDD.
+We only did this for species with at least 10 presences in the plots.
+We only kept models in which at least one of the two bioclimatic indices had a significant effect size (p < 0.05) and for which the value of McFadden's pseudo R^2^ [@RN5066] was higher than 0.15.
+For readers unfamiliar with McFadden's pseudo R^2^, we point out that it tends to have lower values than R^2^ in ordinary least squares regression, and that values between 0.2 and 0.4 represent excellent fit [@RN5067].
+
+To construct scenarios of climatic changes, we used the temporal survey data (10 years, 2009-2018) to describe situations in which the extreme values of today would become the new average.
+To do so, we calculated, for GDD and FDD, the maximum and minimum values recorded in the entire period.
+With the resulting four values, we created four scenarios: hot and snowy (max GDD, min FDD), hot and frozen (max GDD, max FDD), cold and snowy (min GDD, min FDD) and cold and frozen (min GDD, max FDD).
+We used the GLM models to predict the probably of presence for each species and scenario, considering that a probability of 0 in a given scenario would mean the extinction of the species in said scenario.
 
@@
 
+redo environmental fit
+
 Results of the environmental fit:
@@
 .... (Eduardo) (quizás repensar esta parte, pero es importante mantenerla porque es la clave).
+
+# DATA AVAILABILITY
+
+The original data, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/picos.
+A version of record of the repository is deposited in Zenodo.
+
+# LITERATURE CITED
+
+::: {#refs}
+:::
+
+## Figures
+
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Study system."}
+knitr::include_graphics("../results/figures/F1 - Map.png")
+```
+
+```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Study system."}
+knitr::include_graphics("../results/figures/F2 - Temperature logs.png")
+```
+
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Study system."}
+knitr::include_graphics("../results/figures/F3 - Space vs time.png")
+```
+
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Study system."}
+knitr::include_graphics("../results/figures/F4 - NMDS.png")
+```
+
+```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Study system."}
+knitr::include_graphics("../results/figures/F5 - Extinctions.png")
+```
+
+## Tables
+
+```{r echo = FALSE, fig.cap = "**Figure 1** Study system."}
+read.csv("../results/tables/T1 - Summary of the GLM models of species presence.csv", check.names = FALSE) %>%
+  kableExtra::kbl(caption = "Table 1 Summary of the GLM models of species presence") %>%
+  kableExtra::kable_classic(., full_width = F)
+```

### 2022-02-23 src/nmds.R
results first draft ⭐

@@ -51,7 +51,7 @@ header %>%
   filter(Plot %in% sitescores$Plot) %>%
   column_to_rownames(var = "Plot") %>%
   dplyr::select(-Site) %>%
-  select(bio1:GDD) -> env
+  select(FDD:GDD) -> env
 
 env %>% cor() -> biocor
 

### 2022-02-23 src/space-time.R

@@ -6,7 +6,21 @@ bioclim %>%
   gather(Trait, Value, bio1:GDD) %>%
   group_by(Trait, Site, Survey) %>%
   filter(Trait %in% c("GDD", "FDD")) %>%
-  summarise(Delta = max(Value) - min(Value)) %>%
+  summarise(Delta = max(Value) - min(Value)) -> diffs
+
+diffs %>%
+  spread(Survey, Delta) %>%
+  filter(Trait == "FDD") -> fdd
+
+t.test(fdd$Spatial, fdd$Temporal, paired = TRUE, alternative = "greater")
+
+diffs %>%
+  spread(Survey, Delta) %>%
+  filter(Trait == "GDD") -> gdd
+
+t.test(gdd$Spatial, gdd$Temporal, paired = TRUE, alternative = "greater")
+
+diffs %>%
   group_by(Trait, Survey) %>%
   summarise(Mean = mean(Delta), SE = sd(Delta)/sqrt(4)) %>%
   group_by() %>%

### 2022-11-21 doc/manuscript.Rmd

@@
 
-The study was conducted in the central calcareous massif of the Picos de Europa National Park, in northern Spain (**Figure 1A**).
+The study was conducted in the central calcareous massif of the Picos de Europa National Park, in northern Spain (**Fig.
+1A**).
 The study area is a biodiversity hotspot for cold-adapted plants in the Iberian Peninsula and a biogeographical hub for Alpine and Mediterranean lineages in Western Europe (XXX).
@@
 In 2008, we established a long-term ecological research program for monitoring soil climate and vegetation change.
-We selected four study sites along a north-south gradient (**Figure 1B**), reflecting the main macroclimatic variation from Atlantic to Mediterranean influence.
+We selected four study sites along a north-south gradient (**Fig.
+1B**), reflecting the main macroclimatic variation from Atlantic to Mediterranean influence.
 The sites were selected to...
@@
 *Spatial survey.* In 2018, we visited the same areas to study the spatial variation of vegetation and microclimate around the previously sampled areas.
-Using the long-term temperature logger as the central point, we additionally placed 20 iButtons (Thermochron, iButton, Newbury, UK; accuracy: +/- 0.5 ºC from -10 ºC to +65 ºC, resolution: 0.5 ºC, records each 4 hours) in 20 plots of 1 m^2^ separated 10 m from each other in the four cardinal directions (**Figure 1C**).
+Using the long-term temperature logger as the central point, we additionally placed 20 iButtons (Thermochron, iButton, Newbury, UK; accuracy: +/- 0.5 ºC from -10 ºC to +65 ºC, resolution: 0.5 ºC, records each 4 hours) in 20 plots of 1 m^2^ separated 10 m from each other in the four cardinal directions (**Fig.
+1C**).
 At each of the 20 plots, we identified vascular plants and estimated their relative cover in %.
@@
 To do so, we calculated, for GDD and FDD, the maximum and minimum values recorded in the entire period.
-With the resulting four values, we created four scenarios: hot and snowy (max GDD, min FDD), hot and frozen (max GDD, max FDD), cold and snowy (min GDD, min FDD) and cold and frozen (min GDD, max FDD).
+With the resulting four values, we created four scenarios: hot and snowy (max GDD = 2069 ºC, min FDD = 0 ºC), hot and frozen (max GDD = 2069 ºC, max FDD = 247 ºC), cold and snowy (min GDD = 570 ºC, min FDD = 0 ºC) and cold and frozen (min GDD = 570 ºC, max FDD = 247 ºC).
 We used the GLM models to predict the probably of presence for each species and scenario, considering that a probability of 0 in a given scenario would mean the extinction of the species in said scenario.
@@
 
-Along the 10 years of soil temperature monitoring, we found a general trend of increasing temperature, especially in the southwest and warmest site (Hoyo sin Tierra) (Figure 3).
-Two of the sites (Los Cazadores and Los Boches) showed a consistent pattern of continuous snow cover during winter (reflected by temperature records around 0º C).
-In contrast, the two other sites (Hoy Sin Tierri and Hoyo Sin Tierra) showed frost temperatures during most winters (although they may not be constant and intermixed with short snow periods).
-Maximum temperatures were...
-Soil temperature corresponding to the spatial surveys also showed high variation within sites (Figure 3).
-In the four sites we found plots with contrasting patterns of snow cover and frost during winter, reflected by...
-When looking at the whole microclimatic variation among plots (Figure 4), we found that the four sites were represented by a wide range of temperatures, with the exception of one site (Los Cazadores).
-The two main axes of variation were mainly related to FDD and GDD, the latter also correlated with....
-(Eduardo)
+After 10 years of soil temperature monitoring (**Fig.
+2A**), we found that two of the sites (Los Cazadores and Los Boches) showed a consistent pattern of continuous snow cover during winter (i.e.
+snowbed conditions reflected by temperature records around 0 ºC).
+In contrast, the two other sites (Ḥou Sin Tierri and Hoyo Sin Tierra) showed frost temperatures during most winters (i.e.
+fellfield conditions).
+The length of snow cover went from 72 days at Hoyo Sin Tierra to 193 days at Los Boches.
+Mean annual temperature ranged from 3.8 ºC (Los Boches) to 6.2 ºC (Hoyo Sin Tierra).
+The average annual range went from 19.4 ºC (Los Boches) to 23.8 ºC (Hoyo Sin Tierra), and the diurnal range from 2 ºC (Los Boches) to 3 ºC (Ḥou Sin Tierri).
+The absolute maximum was 27.4 ºC, reached at Hoyo Sin Tierra in July 2015, while the absolute minimum was -12 ºC at Ḥou Sin Tierri in December 2016.
+Average annual GDD were lowest at Los Boches (930 ºC) and highest at Hoyo Sin Tierra (1,711 ºC).
+FDD were low at Los Cazadores (3 ºC) and Los Boches (26 ºC), and high at Ḥou Sin Tierri (89 ºC) and Hoyo Sin Tierra (100 ºC).
 
-## Plant diversity and microclimate
+Soil temperature corresponding to the spatial surveys (**Fig.
+2B**) showed high variation within sites.
+In the four sites we found plots representing both snowbed and fellfield conditions.
+Within the plots of a same site, the length of snow cover could go from non-existent (i.e.
+0 days under snow and frost temperatures during the whole year) to as long as 8 months (with a maximum of 234 days, from November to July).
+The annual range of temperatures went from 17.8 ºC to 30.3 ºC, and the diurnal range from 1.6 ºC to 5.5 ºC.
+The absolute maximum was 33 ºC, the absolute minimum -12 ºC.
+GDD ranged from 517 ºC to 1,612 ºC, and FDD from 0 ºC to 206 ºC.
+
+To compare the width of the microclimatic variability in time and space we took, for each site, the difference between the maximum and minimums values of GDD and FDD in time (maxi and min values in 10 years per site) and space (max and min values in 20 plots per site) (**Fig.
+3A**).
+For GDD, the spatial difference was larger than the temporal one, but the difference was only marginally different (paired t-test, one.sided, t = 1.945, df = 3, p-value = 0.073).
+For FDD, the two differences were not significantly different (paired t-test, one.sided, t = -0.29029, df = 3, p-value = 0.6047).
+
+> Quizás sería mejor quitar este último párrafo y la figura 3, y centrarse en los modelos, la parte de time vs.
+space no es nada concluyente.
+Quizás describirla mejor, dentro del estilo de los dos párrafos anteriores.
+
+## Plant diversity
 
@@
 *cantabricum* as separate species), representing % of the total species pool of the study region.
+Of these, 81 species were in the spatial survey plots, and 48 in the temporal survey plots.
 The five most frequent species were *Thymus praecox* subsp.
@@
 
-Changes in species frequency were/were not correlated with XX and XX....
-(Table 1) (Eduardo).
-In the spatial survey (20 plots per site, n = 80; including one plot placed in rocks with no vascular plants) we recorded 81 species, 36 of which were absent from the temporal survey plots.
+## Species responses to microclimate
 
 Changes in species composition (**Fig.
-4**) were associated to GDD (NMDS axis 1) to represent a main gradient from the warmest to the coldest microsites, and to FDD (NMDS2) in a gradient of snow versus frost winters.
-In the compositional space, the sites were poorly differentiated across the four sites (PERMANOVA, adonis) mainly by the effect of one site (Los Boches) with most plots places in the cold and frost margins.
-Etc....
-(Borja).
+4**) were associated to GDD (NMDS axis 1, p < 0.001, R^2^ = 0.778) to represent a main gradient from the warmest to the coldest microsites, and to FDD (NMDS2, p < 0.001, R^2^ = 0.330) in a gradient of snow versus frost microsites In the compositional space, the sites were poorly differentiated along the first axis, although Los Boches occupied the cold and frozen space.
 
-redo environmental fit
-
-Results of the environmental fit:
-
-```{r}
-print(ef1)
-```
-
-Correlations between bioclimatic variables:
-
-```{r}
-print(biocor)
-```
-
-Let us keep GDD, bio7 and FDD for plotting NMDS (low autocorrelation, high r2 in environmental fit)
-
-## Microclimatic ranges in time and space
-
-.... (Eduardo) (quizás repensar esta parte, pero es importante mantenerla porque es la clave).
+Of the 81 species in the spatial plots, 36 had more than 10 occurrences, and we included them in the GLM modeling (full model results in supplementary material **S3**).
+For 16 of these species we produced models that we considered satisfactory (i.e.
+at least one of the two bioclimatic indices had a significant effect size and the value of McFadden's pseudo R^2^ was higher than 0.15) (**Table 1**).
+We found that some species survived only in the hot scenarios (e.g.
+*Androsace villosa*), the cold scenario (e.g.
+*Festuca glacialis*) or the snowy scenarios (e.g.
+*Alchemilla catalaunica*).
+The cold & snowy scenario produced the lowest number of species extinctions and the hot & frozen scenario the most extinctions, with the cold & frozen and the hot & snowy scenarios producing intermediate numbers of extinctions (**Fig.
+5**).
 

### 2023-07-14 src/figure-spp-changes.R
script changes 14 july

@@ -34,20 +34,20 @@ changespp %>%
   xlab(NULL) +
   ylab("Frequency change (%)") +
   ggthemes::theme_tufte() +
-  theme(text = element_text(family = "sans"),
+  theme(text = element_text(family = "sans", size = 20),
         strip.background = element_blank(),
         legend.position = "none", 
         legend.direction = "vertical",
         legend.title = element_blank(),
-        legend.text = element_text(size = 16, face = "italic"), 
+        legend.text = element_text(size = 18, face = "italic"), 
         panel.background = element_rect(color = "black", fill = NULL),
-        strip.text = element_text(size = 12),
-        axis.title = element_text(size = 12),
-        axis.text.x = element_text(size = 12, color = "black"),
-        axis.text.y = element_text(size = 11, color = "black", face = "italic")) -> F4; F4
+        strip.text = element_text(size = 16),
+        axis.title = element_text(size = 18),
+        axis.text.x = element_text(size = 14, color = "black"),
+        axis.text.y = element_text(size = 14, color = "black", face = "italic")) -> F4; F4
 
-ggsave(F4, file = "results/figures/F4.png", 
-       path = NULL, scale = 1, width = 182, height = 145, units = "mm", dpi = 600)
+ggsave(F4, file = "results/figures/clara changes/F4(3).png", 
+       path = NULL, scale = 1, width = 225, height = 145, units = "mm", dpi = 600)
 
 # figure filtered with less than 10 initial observation + images
 library(cowplot)

### 2023-07-14 src/figure-temperature-changes.R

@@ -17,7 +17,8 @@ decomposer <- function(x)
   df
 }
 
-read.csv("data/temporal-survey-temperatures.csv") %>%
+read.csv("data/temporal-survey-temperatures.csv", sep =";") %>%
+  mutate(Time = strptime(as.character(Time), "%d/%m/%Y %H:%M"))%>%
   mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
   group_by(Site) %>%
   do(decomposer(.)) %>%
@@ -25,15 +26,20 @@ read.csv("data/temporal-survey-temperatures.csv") %>%
   mutate(Site = fct_relevel(Site,
                             "Los Cazadores", "Los Boches",
                             "Hou Sin Tierri","Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Hou Sin Tierri" = "Hou Sin Tierri")) -> trends
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> trends
 
 ### Plot trends
+graph_names <- as_labeller (c( "Los Cazadores" ="Los Cazadores \n (Snowbed)", 
+                               "Los Boches" = "Los Boches \n (Snowbed)",
+                               "Ḥou Sin Tierri" = "Ḥou Sin Tierri \n (Fellfield)", 
+                               "Hoyo Sin Tierra" = "Hoyo Sin Tierra \n (Fellfield)"))
 
 trends %>%
   ggplot(aes(Time, Trend)) + 
-  facet_wrap(~ Site, nrow = 1, scales = "free_y") + 
+  facet_wrap(~ Site, nrow = 1, scales = "free_y", labeller = graph_names) + 
   geom_line(color = "firebrick1", alpha = 0.3) +
   geom_smooth(method = "lm", color = "firebrick4") +
+  labs(title = "(A) Temporal trends in soil temperature") +
   xlab("Time (1-h recording inverval)") +
   ylab("Trend (ºC)") +
   ggthemes::theme_tufte() +
@@ -52,7 +58,7 @@ trends %>%
 
 ### Save figure
 
-ggsave(F3A, file = "results/figures/F3A.png", 
+ggsave(F3A, file = "results/figures/clara changes/F3A (2).png", 
        path = NULL, scale = 1, width = 182, height = 182/3, units = "mm", dpi = 600)
[…]

### 2023-07-14 src/glms.R

@@ -374,26 +374,59 @@ models %>%
   ggplot(aes(Scenario, n, fill = Scenario)) +
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_bar(stat = "identity", position = "dodge", color = "black") +
-  labs(title = "(A) 0% Occurrence")+
+  labs(title = "(A) Probability of occurrence = 0")+
   xlab("Scenario") +
   ylab("Local extinctions") +
   ggthemes::theme_tufte() +
   scale_fill_manual(values = c("deepskyblue","slateblue3", "indianred","firebrick")) +
   scale_y_continuous(limits =c (0,10) ,breaks = seq(0, 10, by = 1)) +
-  theme(text = element_text(family = "sans"),
+  theme(text = element_text(family = "sans", size = 20),
         strip.background = element_blank(),
         legend.position = "none", 
         legend.direction = "vertical",
         legend.title = element_blank(),
-        legend.text = element_text(size = 16, face = "italic"), 
+        legend.text = element_text(size = 18, face = "italic"), 
         panel.background = element_rect(color = "black", fill = NULL),
-        strip.text = element_text(size = 12),
-        axis.title = element_text(size = 12),
-        axis.text.x = element_text(size = 8, color = "black"),
-        axis.text.y = element_text(size = 11, color = "black")) ->  F5A; F5A
+        strip.text = element_text(size = 16),
+        axis.title = element_text(size = 18),
+        axis.text.x = element_text(size = 14, color = "black"),
+        axis.text.y = element_text(size = 14, color = "black")) ->  F5A; F5A
 
+models %>%
+  filter(`FDD p.value` < 0.05 | `GDD p.value` < 0.05) %>%
+  filter(rho2 > 0.15) %>%
+  dplyr::select(Taxon, `Cold & Freezing`:`Hot & Snowy`) %>%
+  gather(Scenario, Predict, `Cold & Freezing`:`Hot & Snowy`) %>%
+  #filter(Predict == 0) %>% # 100 % threshold
+  filter(Predict <=5) %>% #95% threshold
+  #mutate(Scenario = factor(Scenario, levels = c("Cold & Snowy", "Cold & Freezing","Hot & Snowy", "Hot & Freezing"))) %>%
+  group_by(Scenario) %>%
+  tally %>%
+  mutate(Scenario = gsub(" & ", "\n", Scenario)) %>%
[…]

### 2023-07-14 src/logs.R

@@ -2,6 +2,10 @@ library(tidyverse); library(urca); library(raster); library(tibbletime); library
 Sys.setlocale("LC_TIME", "English")
 
 ### iButton logs
+graph_names <- as_labeller (c( "Los Cazadores" ="Los Cazadores \n (Snowbed)", 
+                               "Los Boches" = "Los Boches \n (Snowbed)",
+                               "Hou Sin Tierri" = "Hou Sin Tierri \n (Fellfield)", 
+                               "Hoyo Sin Tierra" = "Hoyo Sin Tierra \n (Fellfield)"))
 
 read.csv("data/spatial-survey-temperatures.csv") %>%
   mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
@@ -11,7 +15,7 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
                             "Hou Sin Tierri","Hoyo Sin Tierra")) %>%
   mutate(Site = fct_recode(Site, "Hou Sin Tierri" = "Hou Sin Tierri")) %>%
   ggplot(aes(Time, Temperature, color = Plot)) + 
-  facet_wrap(~ Site, nrow = 1) + 
+  facet_wrap(~ Site, nrow = 1, labeller = graph_names) + 
   geom_hline(yintercept = 0, linetype = "dashed") +
   geom_line() +
   labs(title = "(B) Spatial survey (20 iButtons per site, Oct 2018 - Aug 2019)") +
@@ -29,18 +33,22 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
         strip.text = element_text(size = 12),
         axis.title = element_text(size = 12),
         axis.text.x = element_text(size = 8, color = "black"),
-        axis.text.y = element_text(size = 11, color = "black")) -> F2B
+        axis.text.y = element_text(size = 11, color = "black")) -> F2B;F2B
 
 ### Geoprecision logs
+graph_names <- as_labeller (c( "Los Cazadores" ="Los Cazadores \n (Snowbed)", 
+                               "Los Boches" = "Los Boches \n (Snowbed)",
+                               "Hou Sin Tierri" = "Hou Sin Tierri \n (Fellfield)", 
+                               "Hoyo Sin Tierra" = "Hoyo Sin Tierra \n (Fellfield)"))
 
-read.csv("data/temporal-survey-temperatures.csv") %>% 
+read.csv("data/temporal-survey-temperatures.csv", sep= ";") %>% 
[…]

### 2023-07-14 src/space-time2.R

@@ -2,7 +2,8 @@ library(tidyverse)
 
 ### Temporal survey
 
-read.csv("data/temporal-survey-temperatures.csv") %>%
+read.csv("data/temporal-survey-temperatures.csv", sep = ";") %>%
+  mutate(Time = strptime(as.character(Time), "%d/%m/%Y %H:%M"))%>%
   mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
   mutate(Month = lubridate::month(Time)) %>% 
   mutate(Day = lubridate::day(Time)) %>% 
@@ -16,7 +17,8 @@ read.csv("data/temporal-survey-temperatures.csv") %>%
 
 ### Spatial survey
 
-read.csv("data/spatial-survey-temperatures.csv") %>%
+read.csv("data/spatial-survey-temperatures.csv", sep = ",") %>%
+  mutate(Time = strptime(as.character(Time), "%d/%m/%Y %H:%M"))%>%
   mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
   merge(read.csv("data/spatial-survey-header.csv"))%>%
   mutate(Survey = "Spatial") %>%
@@ -26,6 +28,12 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
 
 rbind(temporal, spatial) -> temps
 
+
+graph_names <- as_labeller (c( "Los Cazadores" ="Los Cazadores \n (Snowbed)", 
+                               "Los Boches" = "Los Boches \n (Snowbed)",
+                               "Hou Sin Tierri" = "Hou Sin Tierri \n (Fellfield)", 
+                               "Hoyo Sin Tierra" = "Hoyo Sin Tierra \n (Fellfield)"))
+
 temps %>%  
   mutate(Site = fct_relevel(Site,
                                      "Los Cazadores", "Los Boches",
@@ -34,7 +42,8 @@ temps %>%
   ggplot(aes(Survey, Temperature, color = Survey, fill = Survey)) + 
   geom_jitter(shape = 16, position = position_jitter(0.15), alpha = 0.1) +
   geom_violin(alpha = 0.05, draw_quantiles = c(0.25, 0.5, 0.75), color = "black") +
-  facet_wrap(~ Site, nrow = 1)  +
+  facet_wrap(~ Site, nrow = 1, labeller = graph_names)  +
+  labs (title = "(B) Spatial vs. Temporal")+
   xlab("Survey") +
   ylab("Temperature (ºC)") +
[…]

### 2023-07-14 src/figure-temperature-changes.R
script changes 14 july

@@ -26,7 +26,7 @@ read.csv("data/temporal-survey-temperatures.csv", sep =";") %>%
   mutate(Site = fct_relevel(Site,
                             "Los Cazadores", "Los Boches",
                             "Hou Sin Tierri","Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> trends
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri"))  -> trends
 
 ### Plot trends
 graph_names <- as_labeller (c( "Los Cazadores" ="Los Cazadores \n (Snowbed)", 

### 2023-07-14 src/logs.R

@@ -66,7 +66,7 @@ read.csv("data/temporal-survey-temperatures.csv", sep= ";") %>%
         strip.text = element_text(size = 12),
         axis.title = element_text(size = 12),
         axis.text.x = element_text(size = 8, color = "black"),
-        axis.text.y = element_text(size = 11, color = "black")) -> F2A;F2A
+        axis.text.y = element_text(size = 11, color = "black")) -> F2A; F2A
 
 ### Combine panels
 

### 2023-07-14 src/map.R

@@ -50,7 +50,7 @@ read.csv("data/temporal-survey-header.csv") %>%
         axis.text = element_blank(),
         axis.ticks = element_blank(),             
         plot.margin = unit(c(0, 0.1, 0, 0), "cm"),
-        axis.title = element_blank()) -> f1a; f1a
+        axis.title = element_blank()) -> f1a;  f1a
 
 ### Panel B - Map of Picos
 

### 2023-07-14 src/nmds.R

@@ -104,7 +104,7 @@ ggplot(sitescores, aes(x = Dim.1, y = Dim.2)) +
   annotate("label", x = -.8, y = 1, label = "Hot & Snowy", fill = "indianred", color = "white", size = 4) +
   annotate("label", x = .7, y = 1, label = "Cold & Snowy", fill = "deepskyblue", color = "white", size = 4) +
   annotate("label", x = -.8, y = -1, label = "Hot & Freezing", fill = "firebrick", color = "white", size = 4) +
-  annotate("label", x = .7, y = -1, label = "Cold & Freezing", fill = "slateblue3", color = "white", size = 4) -> f1d;f1d
+  annotate("label", x = .7, y = -1, label = "Cold & Freezing", fill = "slateblue3", color = "white", size = 4) -> f1d; f1d
 
 ### Save figure
 

### 2023-07-14 src/space-time2.R

@@ -22,7 +22,7 @@ read.csv("data/spatial-survey-temperatures.csv", sep = ",") %>%
   mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
   merge(read.csv("data/spatial-survey-header.csv"))%>%
   mutate(Survey = "Spatial") %>%
-  select(Site, Time, Temperature, Survey) -> spatial
+  select(Site, Time, Temperature, Survey) ->  spatial
 
 ## Bind
 

### 2023-07-14 src/logs.R
revised final draft, figures

@@ -4,7 +4,7 @@ Sys.setlocale("LC_TIME", "English")
 ### iButton logs
 graph_names <- as_labeller (c( "Los Cazadores" ="Los Cazadores \n (Snowbed)", 
                                "Los Boches" = "Los Boches \n (Snowbed)",
-                               "Hou Sin Tierri" = "Hou Sin Tierri \n (Fellfield)", 
+                               "Ḥou Sin Tierri" = "Ḥou Sin Tierri \n (Fellfield)", 
                                "Hoyo Sin Tierra" = "Hoyo Sin Tierra \n (Fellfield)"))
 
 read.csv("data/spatial-survey-temperatures.csv") %>%
@@ -13,7 +13,7 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
   mutate(Site = fct_relevel(Site,
                             "Los Cazadores", "Los Boches",
                             "Hou Sin Tierri","Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Hou Sin Tierri" = "Hou Sin Tierri")) %>%
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) %>%
   ggplot(aes(Time, Temperature, color = Plot)) + 
   facet_wrap(~ Site, nrow = 1, labeller = graph_names) + 
   geom_hline(yintercept = 0, linetype = "dashed") +
@@ -38,15 +38,15 @@ read.csv("data/spatial-survey-temperatures.csv") %>%
 ### Geoprecision logs
 graph_names <- as_labeller (c( "Los Cazadores" ="Los Cazadores \n (Snowbed)", 
                                "Los Boches" = "Los Boches \n (Snowbed)",
-                               "Hou Sin Tierri" = "Hou Sin Tierri \n (Fellfield)", 
+                               "Ḥou Sin Tierri" = "Ḥou Sin Tierri \n (Fellfield)", 
                                "Hoyo Sin Tierra" = "Hoyo Sin Tierra \n (Fellfield)"))
 
 read.csv("data/temporal-survey-temperatures.csv", sep= ";") %>% 
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC", format = "%d/%m/%Y %H:%M")) %>%
   mutate(Site = fct_relevel(Site,
[…]

### 2023-07-14 src/nmds.R

@@ -37,7 +37,7 @@ vegan::scores(nmds, "sites") %>%
   mutate(Site = fct_relevel(Site,
                             "Los Cazadores", "Hou Sin Tierri",
                             "Los Boches", "Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Hou Sin Tierri" = "Hou Sin Tierri")) -> sitescores
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) -> sitescores
 
 ### Species scores
 
@@ -108,7 +108,7 @@ ggplot(sitescores, aes(x = Dim.1, y = Dim.2)) +
 
 ### Save figure
 
-ggsave(f1d, file = "results/figures/F4 - NMDS_cl.png", 
+ggsave(f1d, file = "results/figures/F4 - NMDS_cl (H).png", 
        path = NULL, scale = 1, width = 182, height = 120, units = "mm", dpi = 600)
 # ggsave(fig, file = "results/figures/nmds-species.tiff", device = grDevices::tiff, 
 #        path = NULL, scale = 1, width = 182 height = 182, units = "mm", dpi = 600, compression = "lzw")

### 2023-07-14 src/space-time2.R

@@ -31,14 +31,14 @@ rbind(temporal, spatial) -> temps
 
 graph_names <- as_labeller (c( "Los Cazadores" ="Los Cazadores \n (Snowbed)", 
                                "Los Boches" = "Los Boches \n (Snowbed)",
-                               "Hou Sin Tierri" = "Hou Sin Tierri \n (Fellfield)", 
+                               "Ḥou Sin Tierri" = "Ḥou Sin Tierri \n (Fellfield)", 
                                "Hoyo Sin Tierra" = "Hoyo Sin Tierra \n (Fellfield)"))
 
 temps %>%  
   mutate(Site = fct_relevel(Site,
                                      "Los Cazadores", "Los Boches",
                                      "Hou Sin Tierri", "Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Hou Sin Tierri" = "Hou Sin Tierri")) %>%
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) %>%
   ggplot(aes(Survey, Temperature, color = Survey, fill = Survey)) + 
   geom_jitter(shape = 16, position = position_jitter(0.15), alpha = 0.1) +
   geom_violin(alpha = 0.05, draw_quantiles = c(0.25, 0.5, 0.75), color = "black") +
@@ -79,5 +79,5 @@ ggarrange(F3A, F3B, nrow = 2, heights = c(1.5,3))->F3;F3
 
 ### Save figure
 
-ggsave(F3, file = "results/figures/clara changes/F3 (2).png", 
+ggsave(F3, file = "results/figures/clara changes/F3 (H).png", 
        path = NULL, scale = 1, width = 182, height = 160, units = "mm", dpi = 600)

### 2023-07-14 src/glms con cover.R
clean old versions

@@ -1,103 +1,0 @@
-library(tidyverse)
-
-### Prepare data
-
-read.csv("data/spatial-survey-species.csv") %>%
-  group_by(Taxon) %>%
-  mutate(n = length(Taxon)) %>%
-  filter(n >= 5) %>%
-  mutate(Presence = 1) %>%
-  select(Plot, Taxon, Cover) %>%
-  spread(Taxon, Cover, fill = 0) %>%
-  gather(Taxon, Cover, -Plot) %>%
-  merge(read.csv("results/supplement/S1 - Bioclimatic indices.csv")) -> presences
-
-### Get GLM coefficients
-
-glms <- function(x) {
-  glm(cbind(Cover, 100 - Cover) ~ GDD, family = "binomial", data = x) -> m1
-  broom::tidy(m1)
-}
-
-presences %>%
-  group_by(Taxon) %>%
-  do(glms(.)) %>%
-  gather(Trait, Value, estimate:p.value) %>%
-  mutate(Trait = paste(term, Trait)) %>%
-  select(Taxon, Trait, Value) %>%
-  spread(Trait, Value) %>%
-  select(`Taxon`, 
-         `(Intercept) estimate`, `(Intercept) std.error`, `(Intercept) statistic`, `(Intercept) p.value`, 
-         `GDD estimate`, `GDD std.error`, `GDD statistic`, `GDD p.value`) -> t1
-
-### Get GLM summary statistics
-
-glms.glance <- function(x) {
-  glm(cbind(Cover, 100 - Cover) ~ GDD, family = "binomial", data = x) -> m1
-  broom::glance(m1)
-}
-
-presences %>%
-  group_by(Taxon) %>%
-  do(glms.glance(.)) -> t2
-
-### Get McFadden's pseudo R2 (rho2) https://stats.stackexchange.com/questions/82105/mcfaddens-pseudo-r2-interpretation
-
-glms.mcfadden <- function(x) { 
-  glm(cbind(Cover, 100 - Cover) ~ GDD, family = "binomial", data = x) -> mod
-  glm(cbind(Cover, 100 - Cover) ~ 1, family = "binomial", data = x) -> nullmod
-  data.frame(rho2 = 1 - logLik(mod) / logLik(nullmod)) #https://thestatsgeek.com/2014/02/08/r-squared-in-logistic-regression/
-}
-
-presences %>%
-  group_by(Taxon) %>%
-  do(glms.mcfadden(.)) %>%
[…]

### 2023-07-14 src/logs.R

@@ -74,7 +74,7 @@ cowplot::plot_grid(F2A, F2B, ncol = 1) -> F2; F2
 
 ### Save figure
 
-ggsave(F2, file = "results/figures/clara changes/F2(H).png", 
+ggsave(F2, file = "results/figures/F2.png", 
        path = NULL, scale = 1, width = 182, height = 140, units = "mm", dpi = 600)
 # ggsave(fig, file = "results/figures/temperature-logs.tiff", device = grDevices::tiff, 
 #        path = NULL, scale = 1, width = 182, height = 160, units = "mm", dpi = 600, compression = "lzw")

### 2023-07-14 src/map.R

@@ -2,6 +2,13 @@ library(tidyverse); library(raster); library(tibbletime); library(lubridate); li
 
 ### Raster with elevation for Picos area
 
+# raster::raster("../#data/maps/picos/DEM/DEM.TIF") %>% # Map files are in my home drive
+#   rasterToPoints() %>%
+#   as.data.frame() %>%
+#   rename(lon = x, lat = y, alt = DEM) -> dem 
+
+### Raster with elevation for Picos area
+
 raster::raster("C:/Users/Usuario/OneDrive - Universidad de Oviedo/IMIB/Softwares/GitHub/picos/data/picos/DEM/DEM.TIF") %>% # Map files are in my home drive
   rasterToPoints() %>%
   as.data.frame() %>%
@@ -18,10 +25,14 @@ read.csv("data/temporal-survey-header.csv") %>%
 
 ### Panel A - Map of Europe
 
+# rgdal::readOGR(dsn = "../#data/maps/WWF", 
+#                layer = "wwf_terr_ecos") -> Ecoregions # Map files are in my home drive
 rgdal::readOGR(dsn = "data/maps/WWF", 
                layer = "wwf_terr_ecos") -> Ecoregions # Map files are in my home drive
 rownames(Ecoregions@data) -> Ecoregions@data$id
 fortify(Ecoregions, region = "id") -> Ecoregions.points 
+# plyr::join(Ecoregions.points, Ecoregions@data, by = "id") %>%
+#   inner_join(read.csv("../#data/maps/Biomes.csv"), by = "BIOME") -> land
 plyr::join(Ecoregions.points, Ecoregions@data, by = "id") %>%
   inner_join(read.csv("data/maps/Biomes.csv"), by = "BIOME") -> land
 

### 2023-07-14 src/space-time2.R

@@ -1,83 +1,0 @@
-library(tidyverse)
-
-### Temporal survey
-
-read.csv("data/temporal-survey-temperatures.csv", sep = ";") %>%
-  mutate(Time = strptime(as.character(Time), "%d/%m/%Y %H:%M"))%>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
-  mutate(Month = lubridate::month(Time)) %>% 
-  mutate(Day = lubridate::day(Time)) %>% 
-  mutate(Hour = lubridate::hour(Time)) %>% 
-  filter(Hour %in% c(0, 4, 8, 12, 16, 20)) %>% # Keep same recording hours as ibuttons
-  filter(! Month %in% 9) %>% # Remove missing September days
-  filter(! (Month %in% 8 & Day > 6)) %>% # Remove missing August days
-  filter(! (Month %in% 10 & Day < 3)) %>% # Remove missing October days %>%
-  dplyr::select(-c(Month, Day, Hour)) %>%
-  mutate(Survey = "Temporal") -> temporal
-
-### Spatial survey
-
-read.csv("data/spatial-survey-temperatures.csv", sep = ",") %>%
-  mutate(Time = strptime(as.character(Time), "%d/%m/%Y %H:%M"))%>%
-  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
-  merge(read.csv("data/spatial-survey-header.csv"))%>%
-  mutate(Survey = "Spatial") %>%
-  select(Site, Time, Temperature, Survey) ->  spatial
-
-## Bind
-
-rbind(temporal, spatial) -> temps
-
-
-graph_names <- as_labeller (c( "Los Cazadores" ="Los Cazadores \n (Snowbed)", 
-                               "Los Boches" = "Los Boches \n (Snowbed)",
-                               "Ḥou Sin Tierri" = "Ḥou Sin Tierri \n (Fellfield)", 
-                               "Hoyo Sin Tierra" = "Hoyo Sin Tierra \n (Fellfield)"))
-
-temps %>%  
-  mutate(Site = fct_relevel(Site,
-                                     "Los Cazadores", "Los Boches",
-                                     "Hou Sin Tierri", "Hoyo Sin Tierra")) %>%
-  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) %>%
[…]

### 2023-07-14 README.Rmd
readme updated

@@
 
-# Picos
+# Microclimatic buffering prevents homogenization of relict alpine communities
 
-Proyecto Picos.
+## Contents
 
-## Datos pendientes
+This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.
 
-## Análisis pendientes
+The following materials are available in the folders of this repository:
+
+* `data` Primary data used for the analyses, tables and figures presented in the manuscript.
+* `src` Scripts in `R` language to conduct the analyses and produce the tables and figures presented in the manuscript.
+* `results` Output of the `R` scripts, including figures, tables and supplementary materials.
+* `doc` Manuscript files.
+
+## Citation
+
+Please cite this repository as:
+
+> Jiménez-Alfaro B, Fernández-Pascual E, Espinosa del Alba C, Marcenò C (2023) Microclimatic buffering prevents homogenization of relict alpine communities.

### 2023-07-17 src/glms.R
submitted to JVS

@@ -10,7 +10,7 @@ read.csv("data/spatial-survey-species.csv") %>%
   select(Plot, Taxon, Presence) %>%
   spread(Taxon, Presence, fill = 0) %>%
   gather(Taxon, Presence, -Plot) %>%
-  merge(read.csv("results/supplement/S1 - Bioclimatic indices.csv")) -> presences
+  merge(read.csv("results/supplement/Table S1 - Bioclimatic indices.csv", sep = ";")) -> presences
 
 ### Get GLM coefficients
 
@@ -58,12 +58,12 @@ presences %>%
 ### Scenarios
 
 merge(
-read.csv("results/supplement/S1 - Bioclimatic indices.csv") %>%
+  read.csv("results/supplement/Table S1 - Bioclimatic indices.csv", sep = ";") %>%
   select(Survey, GDD) %>%
   filter(Survey == "Temporal") %>%
   # group_by(Survey) %>%
   filter(GDD == max(GDD) | GDD == min(GDD)),
-read.csv("results/supplement/S1 - Bioclimatic indices.csv") %>%
+  read.csv("results/supplement/Table S1 - Bioclimatic indices.csv", sep = ";") %>%
   select(Survey, FDD) %>%
   filter(Survey == "Temporal") %>%
   # group_by(Survey) %>%
@@ -380,15 +380,15 @@ models %>%
   ggthemes::theme_tufte() +
   scale_fill_manual(values = c("deepskyblue","slateblue3", "indianred","firebrick")) +
   scale_y_continuous(limits =c (0,10) ,breaks = seq(0, 10, by = 1)) +
-  theme(text = element_text(family = "sans", size = 20),
+  theme(text = element_text(family = "sans", size = 12),
         strip.background = element_blank(),
         legend.position = "none", 
         legend.direction = "vertical",
         legend.title = element_blank(),
-        legend.text = element_text(size = 18, face = "italic"), 
+        legend.text = element_text(size = 14, face = "italic"), 
         panel.background = element_rect(color = "black", fill = NULL),
-        strip.text = element_text(size = 16),
-        axis.title = element_text(size = 18),
+        strip.text = element_text(size = 14),
+        axis.title = element_text(size = 14),
         axis.text.x = element_text(size = 14, color = "black"),
         axis.text.y = element_text(size = 14, color = "black")) ->  F5A; F5A
 
@@ -413,20 +413,20 @@ models %>%
   ggthemes::theme_tufte() +
[…]

### 2023-07-17 src/space-time2.R

@@ -1,0 +1,83 @@
+library(tidyverse)
+
+### Temporal survey
+
+read.csv("data/temporal-survey-temperatures.csv", sep = ";") %>%
+  mutate(Time = strptime(as.character(Time), "%d/%m/%Y %H:%M"))%>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>% 
+  mutate(Month = lubridate::month(Time)) %>% 
+  mutate(Day = lubridate::day(Time)) %>% 
+  mutate(Hour = lubridate::hour(Time)) %>% 
+  filter(Hour %in% c(0, 4, 8, 12, 16, 20)) %>% # Keep same recording hours as ibuttons
+  filter(! Month %in% 9) %>% # Remove missing September days
+  filter(! (Month %in% 8 & Day > 6)) %>% # Remove missing August days
+  filter(! (Month %in% 10 & Day < 3)) %>% # Remove missing October days %>%
+  dplyr::select(-c(Month, Day, Hour)) %>%
+  mutate(Survey = "Temporal") -> temporal
+
+### Spatial survey
+
+read.csv("data/spatial-survey-temperatures.csv", sep = ",") %>%
+  mutate(Time = strptime(as.character(Time), "%d/%m/%Y %H:%M"))%>%
+  mutate(Time = as.POSIXct(Time, tz = "UTC")) %>%
+  merge(read.csv("data/spatial-survey-header.csv"))%>%
+  mutate(Survey = "Spatial") %>%
+  select(Site, Time, Temperature, Survey) ->  spatial
+
+## Bind
+
+rbind(temporal, spatial) -> temps
+
+
+graph_names <- as_labeller (c( "Los Cazadores" ="Los Cazadores \n (Snowbed)", 
+                               "Los Boches" = "Los Boches \n (Snowbed)",
+                               "Ḥou Sin Tierri" = "Hou Sin Tierri \n (Fellfield)", 
+                               "Hoyo Sin Tierra" = "Hoyo Sin Tierra \n (Fellfield)"))
+
+temps %>%  
+  mutate(Site = fct_relevel(Site,
+                                     "Los Cazadores", "Los Boches",
+                                     "Hou Sin Tierri", "Hoyo Sin Tierra")) %>%
+  mutate(Site = fct_recode(Site, "Ḥou Sin Tierri" = "Hou Sin Tierri")) %>%
[…]

### 2024-01-30 README.Rmd
title

@@
 
-# Microclimatic buffering prevents homogenization of relict alpine communities
+# Spatiotemporal patterns of microclimatic buffering in relict alpine communities
 
@@
 
-> Jiménez-Alfaro B, Fernández-Pascual E, Espinosa del Alba C, Marcenò C (2023) Microclimatic buffering prevents homogenization of relict alpine communities.
+> Jiménez-Alfaro B, Fernández-Pascual E, Espinosa del Alba C, Marcenò C (2023) Spatiotemporal patterns of microclimatic buffering in relict alpine communities.
